### 2022-12-13 research/notes.tex
~~

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{pmboxdraw}
\usepackage{minted}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}

\begin{document}
\section{Introduction}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

Transcranial magnetic stimulation (TMS) is a common, non-invasive experimental technique used to evoke action potentials in cortical regions of the brain. In particular, researchers often target the motor cortex and measure the motor evoked potential (MEP) aroused by the stimulation. When the purpose of the experiment is to inquire on neuroplasticity, such stimulations are performed under a \textit{paired-pulse paradigm}.

The \textit{paired-pulse paradigm} (or \textit{double pulse paradigm}) consists in eliciting a series of two temporally proximate pulses (in the order of miliseconds). The evoked potentials of the double-stimulations are compared to those of single test stimulations, and their relative amplitude is taken as a proxy of neuroplasticity in the brain. The time separating each of the paired-pulses is termed \textit{interstimulous interval} (ISI). It is the general case that low intervals (4 or 5 miliseconds) produce intracortical inhibition, with the evoked potentials of paired stimulations being generally lower than those of single pulses. Greater intervals, on the other hand, tend to produce facilitation.

In the context of this paper, we shall term any coefficient that serves to represent the proportional relationship between the potentials evoked by paired and test pulses a \textit{measure of relative amplitude}. Measures of relative amplitude in neuroscience are generally computed at the subject level. This is reasonable, since hypotheses generally deal with differences across subject groups. However, TMS experimental results are pulse-specific evoked potentials, and transforming them to subject- or group-specific measures implies lowering the data resolution (i.e. representing the same phenomena with less information). In times when scientific inquiry can greatly benefit from forms of artificial intelligence that rely on large amounts of data, it is limiting to count only with low-resolution measures.

Hence, the goal of this paper is to provide pulse-specific measures of relative amplitude. We formulate coefficients that represent the relative amplitude of individual paired-pulses, keeping data resolution at its highest. Thus, from a computational perspective, our objective is constrained to the sphere of feature engineering. This allows for otherwise unfeasible computational analyses of TMS data, such as the use of machine learning models for the detection of different pulse-response patterns among different groups of clinical subjects.

We show such features allow for a machine learning classifier to correctly determine whether a subject belongs to one of four clinical categories based only on its evoked potentials across different inter-stimulous intervals with an accuracy of up to $90\%$.

\section{Relative amplitude features}

For simplicity, we will deal with the hypothetical situation in which a single ISI was used for paired stimulations. All of our results generalize to different ISI.

\subsection{Definitions}

Let $k$ be the number of experimental subjects in some subject group, to each of whom $n$ paired stimulations and $m$ test stimulations were elicited.

\begin{definition}

    Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices representing the paired and test potentials evoked across each of the $k$ subjects, such that

    \begin{equation}
    \textbf{P} := \begin{bmatrix}x_{11} & x_{12}& \cdots & x_{1k} \\
    x_{21} & x_{22}& \cdots & x_{2k} \\
    & & ... & \\
    x_{n1} & x_{n2}& \cdots & x_{nk} 
    \end{bmatrix} ~~
    \textbf{T}:= \begin{bmatrix}t_{11} & t_{12}& ... & t_{1k} \\
    t_{21} & t_{22}& \cdots & t_{2k} \\
    & & ... & \\
    t_{m1} & t_{m2}& \cdots & t_{mk} 
    \end{bmatrix}
    \end{equation}
\end{definition}

\begin{definition}
Let $\mu_t := \begin{bmatrix} \mu_1, \mu_2, \ldots, \mu_k \end{bmatrix}^\top$ be a vector where $\mu_i$ is the average test potential for the $i$th subject. In other words,

\begin{equation}
    \mu_i = \frac{1}{m}\sum_{j=1}^m t_{ji}
\end{equation}
\end{definition}

\begin{remark}
    Evoked potentials of both paired and test pulses conform to a Gamma distribution, as discussed in the appendix (?). This means it is always assumed $\forall x \in \textbf{P}, \forall t\in \textbf{T}|x, t \in \mathbb{R}^+$.
\end{remark}

\subsection{The $\rho$ and $\delta$ features}

\begin{definition}
Let $x \in \textbf{P}_{\star i}$ be a single paired-stimulation MEP in the $i$th experimental subject, and $\textbf{t} = \textbf{T}_{\star i}$ the vector containing all test MEPs of that subject. Then we define two pulse-specific relative amplitude measures,

\begin{equation}
    \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
\end{equation}

\begin{equation}
    \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j}
\end{equation}
\end{definition}

\begin{remark}
    $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$. (For a proof of this property, consult the appendix.)
\end{remark}

Notice that $\rho$ is the proportion between the potential $x$, evoked by a paired stimulation, with respect to the average potential of single test stimulations. On the other hand, $\delta$ is the average proportion of $x$ with respect to each single test pulse.

The $\rho$ function is not an entirely new contribution. The traditional group level relative amplitude measure is conceived as the average, across all subjects in a group, of the average paired response divided by the average test response of each subject. If we let $S_i$ be the average relative amplitude of the $i$th subject, then $S_i$ has ben traditionally defined as

\begin{equation}
    S_i=\frac{\frac{1}{n}\sum_{j=1}^n x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
\end{equation}

as it is easy to see from decomposing the sum in the numerator. In other words, the traditionally used measure of relative amplitude, at the subject level, has always been the average $\rho$ in a subject.

\subsection{The weighted variants $\rho_w, \delta_w$}

As stated earlier, action potentials evoked by trasncranial magnetic stimulation follow a Gamma distribution; one that closely resembles an exponential distribution (see \textbf{Statistics} section). A valid form of data augmentation that also attempts to deal with the presence of outliers is to weight the averages involved in $\rho, \delta$ using inverse-variance weights. The attempt is to allow for potentials that lay at the tail of the distribution to contribute in a way proportional to their probability.

If we let $\rho_w, \delta_w$ be the weighted versions of $\rho, \delta$, then we have

\begin{align}
    \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\
    \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m w_j}
\end{align}

where $\textbf{w}$ is some appropriate weight vector. Since the purpose of $\textbf{w}$ is to reduce the contribution of outliers to the overall measure of relative amplitude, it is to be expected that the distribution of $\textbf{w}$ closely resembles that of evoked potentials.

\section{Empirical results}

We used data collected across $N = ?$ subjects at the \textit{Laboratory for the Study of Slow-wave sleep}, University of Pennsylvania, with $H = ?$ healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD). Transcranial magnetic stimulation of the motor cortex was elicited to them after a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep. Motor evoked potentials were measured via an electrode (?) placed on the subjects' thumb. In the slow-wave disruption session, an auditory stimulus with sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet not strong enough to wake the subjects, was elicited. This experimental setting produces four distinct categories, two depending on the subject group and two on the type of sleep session underwent.

We trained the exact same random forests model first on the data without pulse-specific measures of relative amplitude, then on the data including some or all the pulse-specific measures of relative amplitude. The model's task was to classifiy each TMS stimulation into one of the four experimental categories. In other words, it faced the challenge of establishing to which type of subject, and under which type of session, each stimulation was elicited, based only on the evoked potential, the ISI and ---later--- the relative amplitude measures of the stimulation. Our objective was to evaluate whether the inclusion of pulse-specific measures of relative amplitude improved the model's accuracy, and if so in what degree.

The exact line of code that defined the model and its parameters was

\begin{minted}{python}
    bag_clf = AdaBoostClassifier(
        DecisionTreeClassifier(max_depth=8), 
        n_estimators=20, 
        algorithm='SAMME.R', 
        learning_rate=0.01)
\end{minted}

\subsection{Raw data}

When trained on the raw data, without the inclusion of pulse-specific relative amplitude measures, the model's accuracy was $A\approx34.2 \%$.

\section{Appendix}

\subsection*{Proofs}

\textbf{Theorem: } $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$

Recall that

\begin{align} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\ \rho(x) &= \frac{xm}{\sum^m t_j} \end{align}

Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m t_j$. We operate under the assumption that $t_i \in \mathbb{R}^+$. It is the case that

\begin{align} \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\
\iff S_2^mS_1^m &\geq m^2
\end{align}

This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So we may assume $S^k_1 S^k_2 \geq k^2$ . We may only show now that

\begin{equation}
S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
\end{equation}

\begin{align} S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\
(S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2+1 &\geq k^2+2k+1 \\
S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k
\end{align}

We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k + \frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that

\begin{align}\frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
\iff\frac{1}{t_{k+1}}(t_1+t_2+...+t_k)+t_{k+1}(\frac{1}{t_1}+\frac{1}{t_2}+...+\frac{1}{t_k}) &\geq 2k \\
\iff \Big(\frac{t_1}{t_{k+1}}+\frac{t_2}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+\frac{t_{k+1}}{t_2}+...+\frac{t_{k+1}}{t_k}\Big) &\geq 2k \\
\iff \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... + n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
\end{align}

We have $\min f=2$  for $f(x)=x+\frac{1}{x}$ for $x \in \mathbb{R}^+$. Then $\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$ for $a,..., n \in \mathbb{R}^+$, which concludes the demonstration.

\end{document}

### 2022-12-13 R/loader.R

@@ -1,4 +1,3 @@
-getwd()
 library(tidyr)
 library(readr)
 library(dplyr)
@@ -16,6 +15,7 @@ ISI_ORDER_120 <- c(
     0, -1, 15, 0, 15, 0, 4, 0, 8, 0, 20, 0, 4, 0, 20,
     0, 4, 0, -1, -1, 10, 0, 20, 0, -1, 15, 0, 5, 0, 20, 0, 4, 0
 )
+
 ISI_ORDER_72 <- c(
     10, 8, -1, 10, -1, 20, -1, 15, 10, 20, 8, 10, 15, -1, 8, -1, 10, -1,
     5, -1, 5, 10, 8, 4, 8, -1, 4, -1, 4, 5, -1, 20, -1, -1, -1, 5, 5,
@@ -27,9 +27,6 @@ WD <- getwd()
 DATA_DIR <- paste(WD, "Data/", sep = "/")
 PARTS_INFO <- read_csv("TMS_ptinfo.csv")
 PARTS_INFO
-
-
-
 
 #' Use .txt file name to find subject number and return it as numeric.
 #'
@@ -71,7 +68,8 @@ get_session_as_number <- function(session_name) {
     )
 }
 
-
+# Horrible but serves its purpose and will
+# only be called once in human history.
 get_label <- function(df) {
     if ((df$Group[1] == 1) && (df$Type == "SWD")) {
         return("HC SWD")
@@ -90,6 +88,7 @@ get_label <- function(df) {
 load_data <- function() {
     df <- tibble()
     for (file in list.files(DATA_DIR)) {
+        print(file)
         path <- paste(DATA_DIR, file, sep = "/")
 
         t <- read_delim(path,
@@ -150,6 +149,7 @@ recode_by_patterns <- function(vec, pattern, subpatterns, new_value) {
     return(vec)
 }
 
+#' Wrapper
 get_subject_data <- function(df, subject) {
     df <- df %>% subset(Subject == subject)
     return(df)
@@ -165,7 +165,9 @@ create_isi_col <- function(df) {
         } else if (len == 72) # If this ICF has 120 pulses.
         {
             vec <- ISI_ORDER_72
-        } else {
+        } else if (len == 118) {
+        vec <- ISI_ORDER_120[1:118] # Special case (see Elly's mails)
+    } else {
         warning(paste("ICF INTEGRITY WARNING: ", len, " on subject ", df$Subject[1]))
         return
     }
[…]

### 2022-12-13 research/jl/analysis.jl

@@ -1,0 +1,34 @@
+using DataFrames
+using CSV
+using Statistics
+include("math.jl")
+
+function sra(pulses, subject, session)
+    """Computes standard relative amplitude given
+    a subject's d vector on a given session."""
+    # Using global variable
+    t = get_test_pulses(subject, session)
+    mean(pulses) / mean(t)
+end
+
+function get_test_pulses(subject, session)
+    """Wrapper to avoid verbose expression"""
+    filter([:ISI, :Subject, :Session] => (i, x, y) -> i == -1 && x == subject && y == session, df).EMGPeakToPeak
+end
+
+function subject_analysis(df)
+    """Performs subject-level ICF/ICI analysis by computing the standard 
+    relative amplitude on each session-subject subgroup per each ISI."""
+    df = filter(:ISI => x -> x != 0 && x != -1, df)
+    gdf = groupby(df, [:ISI, :Subject, :Session, :Label])
+    an = combine(gdf, [:EMGPeakToPeak, :Subject, :Session] => (x, y, z) -> sra(x, y[1], z[1]))
+    rename!(an, 5 => :SRA)
+    sort(an, [:Subject, :Session, :ISI])
+end
+
+function group_analysis(an, col=:SRA)
+    """Computes mean relative amplitude per ISI on each subject group."""
+    an = groupby(an, [:Label, :ISI])
+    an = combine(an, [col] => x -> mean(x))
+    rename!(an, 3 => :SRA)
+end

### 2022-12-13 research/jl/math.jl

@@ -1,0 +1,53 @@
+
+function weighted_avg(a, penalties)
+    sum(a .* penalties) / sum(penalties)
+end
+
+function p(d::Number, t::Vector, f=Φ)::Number
+    """Computes weighted, pulse-specific relative
+    amplitude given a paired-pulse d and a series
+    of test pulses t. Weights are computed via some 
+    penalty f."""
+    w = broadcast(f, rzs(t))
+    p = (d ./ t) .* w
+    sum(p) / sum(w)
+end
+
+function p(d::Vector, t::Vector, f=Φ)::Vector
+    h(x) = p(x, t, f)
+    broadcast(h, d)
+end
+
+function p(d::SubArray, t::Vector, f=Φ)::Vector
+    h(x) = p(x, t, f)
+    broadcast(h, d)
+end
+
+function ℒ(θ::Vector, d::Vector, f)
+    w = broadcast(f, d)
+    sum(θ .* w) / sum(w)
+end
+
+# Penalty functions. See fourth chapter
+# of paper in progress.
+
+function Φ(x, α=3)
+    α / (x^2 + α)
+end
+
+function Ψ(x, α=1)
+    exp(-(x^2 / α))
+end
+
+function ∅(x)
+    x
+end
+
+# Robust z-score
+
+function rzs(x::Vector)::Vector
+    """Given a vector x, returns a vector with the 
+    robust z-scores of each element of x."""
+    mad = median(abs.(x .- median(x)))
+    0.675 * (x .- median(x)) / mad
+end

### 2023-04-12 research/v2.tex
Gitignore and resarch LaTex

--- 2022-12-13 research/notes.tex
@@
 \usepackage{pmboxdraw}
-\usepackage{minted}
-\usepackage{graphicx}
-\graphicspath{ {./images/} }
+\usepackage{graphicx} \graphicspath{
+{./images/} }
+\usepackage{indentfirst}
+\usepackage{multicol}
+\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
 \DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}
+\newtheorem{theorem}{Theorem}
 
 \begin{document}
+
+\begin{multicols}{2}
 \section{Introduction}
@@
 
-Transcranial magnetic stimulation (TMS) is a common, non-invasive experimental technique used to evoke action potentials in cortical regions of the brain.
-In particular, researchers often target the motor cortex and measure the motor evoked potential (MEP) aroused by the stimulation.
-When the purpose of the experiment is to inquire on neuroplasticity, such stimulations are performed under a \textit{paired-pulse paradigm}.
+Transcranial magnetic stimulation (TMS) is a common,
+non-invasive experimental technique used to evoke action
+potentials in cortical regions of the brain. In particular,
+researchers often target the motor cortex and measure the
+motor evoked potential (MEP) aroused by the stimulation.
+When the purpose of the experiment is to inquire on
+neuroplasticity, such stimulations are performed under the
+\textit{paired pulse paradigm}.
 
-The \textit{paired-pulse paradigm} (or \textit{double pulse paradigm}) consists in eliciting a series of two temporally proximate pulses (in the order of miliseconds).
-The evoked potentials of the double-stimulations are compared to those of single test stimulations, and their relative amplitude is taken as a proxy of neuroplasticity in the brain.
-The time separating each of the paired-pulses is termed \textit{interstimulous interval} (ISI).
-It is the general case that low intervals (4 or 5 miliseconds) produce intracortical inhibition, with the evoked potentials of paired stimulations being generally lower than those of single pulses.
-Greater intervals, on the other hand, tend to produce facilitation.
+The \textit{paired pulse paradigm} (or \textit{double pulse
+paradigm}) consists in eliciting a series of two temporally
+proximate pulses (in the order of miliseconds). The evoked
+potentials of the double-stimulation are compared to those
+of single test stimulations, and their relative amplitude is
+taken as a proxy of neuroplasticity in the brain. The time
+separating each of the paired pulses is termed
+\textit{interstimulus interval} (ISI). It is the general
+case that low intervals (4 or 5 miliseconds) produce
+intracortical inhibition, with the evoked potentials of
+paired stimulations being generally lower than those of
+single pulses. Greater intervals, on the other hand, tend to
+produce facilitation.
 
-In the context of this paper, we shall term any coefficient that serves to represent the proportional relationship between the potentials evoked by paired and test pulses a \textit{measure of relative amplitude}.
-Measures of relative amplitude in neuroscience are generally computed at the subject level.
-This is reasonable, since hypotheses generally deal with differences across subject groups.
-However, TMS experimental results are pulse-specific evoked potentials, and transforming them to subject- or group-specific measures implies lowering the data resolution (i.e.
-representing the same phenomena with less information).
-In times when scientific inquiry can greatly benefit from forms of artificial intelligence that rely on large amounts of data, it is limiting to count only with low-resolution measures.
+In the context of this paper, we shall term any coefficient
+that serves to represent the proportional relationship
+between the potentials evoked by paired and test pulses a
+\textit{measure of relative amplitude}. Measures of relative
+amplitude in neuroscience are generally computed at the
+subject level. This is reasonable, since hypotheses
+generally deal with differences across subject groups.
+However, TMS experimental results are pulse-specific evoked
+potentials, and transforming them to subject- or
+group-specific measures implies down-scaling the data resolution.
 
-Hence, the goal of this paper is to provide pulse-specific measures of relative amplitude.
-We formulate coefficients that represent the relative amplitude of individual paired-pulses, keeping data resolution at its highest.
-Thus, from a computational perspective, our objective is constrained to the sphere of feature engineering.
-This allows for otherwise unfeasible computational analyses of TMS data, such as the use of machine learning models for the detection of different pulse-response patterns among different groups of clinical subjects.
+The goal of this paper is to provide pulse-specific measures of relative
+amplitude. This is, coefficients that represent the relative amplitude of each
+individual paired pulse with respect to the set of test pulses in an
+experimental session. The purpose of this endeavor is to keep data resolution
+at its highest, which on its turn allows for the implementation of data-driven
+artificial intelligence in the analysis of the 
+experimental results. Thus, from a computer science perspective, our objective is
+constrained to the sphere of feature engineering.
 
-We show such features allow for a machine learning classifier to correctly determine whether a subject belongs to one of four clinical categories based only on its evoked potentials across different inter-stimulous intervals with an accuracy of up to $90\%$.
+We will show how pulse-specific measures of relative amplitude allow for
+otherwise unfeasible computational analyses of TMS data, such as the use of
+machine learning models for the detection of different pulse-response patterns
+among different groups of clinical subjects. In particular, we will show they
+allow for a machine learning classifier to correctly determine whether a
+subject belongs to one of four clinical categories, based only on its evoked
+potentials and across different inter-stimulus intervals, with an accuracy of up to
+$90\%$.
 
@@
 
-For simplicity, we will deal with the hypothetical situation in which a single ISI was used for paired stimulations. All of our results generalize to different ISI.
+For simplicity, we will deal with the hypothetical situation
+in which a single ISI was used for paired stimulations. All
+of our results generalize to different inter-stimulus intervals.
 
@@
 
-Let $k$ be the number of experimental subjects in some subject group, to each of whom $n$ paired stimulations and $m$ test stimulations were elicited.
+Let $k$ be the number of experimental subjects in some
+subject group $\mathcal{G}$, to each of whom $n$ paired stimulations and
+$m$ test stimulations were elicited.
 
@@
 
-    Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices representing the paired and test potentials evoked across each of the $k$ subjects, such that
+Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices representing
+the paired and test potentials evoked across each of the $k$ subjects, such that
 
-    \begin{equation}
-    \textbf{P} := \begin{bmatrix}x_{11} & x_{12}& \cdots & x_{1k} \\
-    x_{21} & x_{22}& \cdots & x_{2k} \\
-    & & ... & \\
-    x_{n1} & x_{n2}& \cdots & x_{nk} 
-    \end{bmatrix} ~~
-    \textbf{T}:= \begin{bmatrix}t_{11} & t_{12}& ... & t_{1k} \\
-    t_{21} & t_{22}& \cdots & t_{2k} \\
-    & & ... & \\
-    t_{m1} & t_{m2}& \cdots & t_{mk} 
-    \end{bmatrix}
-    \end{equation}
-\end{definition}
+\begin{equation*} \textbf{P} := \begin{bmatrix}x_{11} &
+    x_{12}& \ldots & x_{1k} \\ x_{21} & x_{22}& \ldots &
+    x_{2k} \\ \vdots & \vdots &\ddots& \vdots \\ x_{n1} & x_{n2}& \cdots & x_{nk}
+\end{bmatrix} ~~ \textbf{T}:= \begin{bmatrix}t_{11} &
+t_{12}& \ldots & t_{1k} \\ t_{21} & t_{22}& \cdots & t_{2k} \\
+      \vdots & \vdots &\ddots& \vdots \\ t_{m1} & t_{m2}& \cdots & t_{mk}
+\end{bmatrix} \end{equation*} \end{definition}
 
 \begin{definition} 
-Let $\mu_t := \begin{bmatrix} \mu_1, \mu_2, \ldots, \mu_k \end{bmatrix}^\top$ be a vector where $\mu_i$ is the average test potential for the $i$th subject. In other words,
+    Let $\mu_t := \begin{bmatrix} \mu_1 & \mu_2 & \ldots & \mu_k
+    \end{bmatrix}^\top$ be such that $\mu_i$ is the average test potential for
+    the $i$th subject. In other words,
 
@@
     \mu_i = \frac{1}{m}\sum_{j=1}^m t_{ji}
-\end{equation}
-\end{definition}
+\end{equation} \end{definition}
 
 \begin{remark} 
-    Evoked potentials of both paired and test pulses conform to a Gamma distribution, as discussed in the appendix (?).
-This means it is always assumed $\forall x \in \textbf{P}, \forall t\in \textbf{T}|x, t \in \mathbb{R}^+$.
+    Evoked potentials of both paired and test pulses conform to a Gamma
+    distribution, as discussed in the appendix (?). This means it is always an
+    implicit assumption that $\langle\forall x, t : x \in \textbf{P}, t\in \textbf{T}
+    : x, t \in \mathbb{R}^+\rangle$. 
 \end{remark}
@@
 \begin{definition} 
-Let $x \in \textbf{P}_{\star i}$ be a single paired-stimulation MEP in the $i$th experimental subject, and $\textbf{t} = \textbf{T}_{\star i}$ the vector containing all test MEPs of that subject.
-Then we define two pulse-specific relative amplitude measures,
+    Let $x \in \textbf{P}_{\star i}$ be the MEP of a single paired stimulation
+    elicited on the $i$th experimental subject, and $\textbf{t} =
+    \textbf{T}_{\star i}$ the vector containing the MEPs of all test pulses
+    elicited on that subject. Then we define two pulse-specific relative
+    amplitude measures,
 
@@
 \begin{remark} 
-    $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$. (For a proof of this property, consult the appendix.)
+    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
+    property, consult the appendix.) 
 \end{remark}
 
-Notice that $\rho$ is the proportion between the potential $x$, evoked by a paired stimulation, with respect to the average potential of single test stimulations.
-On the other hand, $\delta$ is the average proportion of $x$ with respect to each single test pulse.
+Notice that $\rho(x)$ is the proportion between the potential
+$x$, evoked by a paired stimulation, with respect to the
+average potential of single test stimulations. On the other
+hand, $\delta(x)$ is the average proportion of $x$ with respect
+to each single test pulse.
 
-The $\rho$ function is not an entirely new contribution.
-The traditional group level relative amplitude measure is conceived as the average, across all subjects in a group, of the average paired response divided by the average test response of each subject.
-If we let $S_i$ be the average relative amplitude of the $i$th subject, then $S_i$ has ben traditionally defined as
+The $\rho$ function is not an entirely new contribution. Traditionally, the
+group-level relative amplitude measure is conceived as the average, across all
+subjects in a group, of the average paired response divided by the average test
+response. This means that, if we let $S_i$ be the average relative amplitude of
+the $i$th subject, then $S_i$ has been traditionally defined as
 
 \begin{equation} 
-    S_i=\frac{\frac{1}{n}\sum_{j=1}^n x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
+    S_i=\frac{\frac{1}{n}\sum_{j=1}^n x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} = 
+    \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n} 
 \end{equation}
 
-as it is easy to see from decomposing the sum in the numerator.
-In other words, the traditionally used measure of relative amplitude, at the subject level, has always been the average $\rho$ in a subject.
+as it is easy to see from decomposing the sum in the
+numerator. In other words, the traditionally used measure of
+relative amplitude, at the subject level, has always been
+the average $\rho$ in a subject, albeit it was never defined explicitly.
 
@@
 
-As stated earlier, action potentials evoked by trasncranial magnetic stimulation follow a Gamma distribution; one that closely resembles an exponential distribution (see \textbf{Statistics} section).
-A valid form of data augmentation that also attempts to deal with the presence of outliers is to weight the averages involved in $\rho, \delta$ using inverse-variance weights.
-The attempt is to allow for potentials that lay at the tail of the distribution to contribute in a way proportional to their probability.
+As stated earlier, action potentials evoked by transcranial
+magnetic stimulations follow a Gamma distribution; one that
+closely resembles an exponential distribution (see
+\textbf{Statistics} section). A valid form of data
+augmentation that also attempts to deal with the presence of
+outliers is to weight the averages involved in $\rho,
+\delta$ using inverse-variance weights. The attempt is to
+allow for potentials that lay at the tail of the
+distribution to contribute in a way proportional to their
+probability.
 
-If we let $\rho_w, \delta_w$ be the weighted versions of $\rho, \delta$, then we have
+If we let $\rho_w, \delta_w$ be the weighted versions of
+$\rho, \delta$, then we have
 
-\begin{align}
-    \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\
-    \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m w_j}
-\end{align}
+        \begin{align} \rho_w(x) &:= \frac{xm\sum_{j=1}^m
+        w_j}{\sum_{j=1}^m t_j w_j} \\ \delta_w(x) &:=
+    \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m
+w_j} \end{align}
 
-where $\textbf{w}$ is some appropriate weight vector.
-Since the purpose of $\textbf{w}$ is to reduce the contribution of outliers to the overall measure of relative amplitude, it is to be expected that the distribution of $\textbf{w}$ closely resembles that of evoked potentials.
+where $\textbf{w}$ is some appropriate weight vector. Since
+the purpose of $\textbf{w}$ is to reduce the contribution of
+outliers to the overall measure of relative amplitude, it is
+to be expected that the distribution of $\textbf{w}$ closely
+resembles that of evoked potentials.
 
@@
 
-We used data collected across $N = ?$ subjects at the \textit{Laboratory for the Study of Slow-wave sleep}, University of Pennsylvania, with $H = ?$ healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
-Transcranial magnetic stimulation of the motor cortex was elicited to them after a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
-Motor evoked potentials were measured via an electrode (?) placed on the subjects' thumb.
-In the slow-wave disruption session, an auditory stimulus with sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet not strong enough to wake the subjects, was elicited.
-This experimental setting produces four distinct categories, two depending on the subject group and two on the type of sleep session underwent.
+We used data collected across $N = ?$ subjects at the \textit{Laboratory for the
+Study of Slow-wave sleep activity}, University of Pennsylvania, with $H = ?$
+healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
+Transcranial magnetic stimulation of the motor cortex was elicited to them after
+a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
+Motor evoked potentials were measured via an electrode (?) placed on the
+subjects' thumb. In the slow-wave disruption session, an auditory stimulus with
+sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet
+not strong enough to wake the subjects, was elicited. This experimental setting
+produces four distinct categories, two depending on the subject group and two on
+the type of sleep session underwent.
 
-We trained the exact same random forests model first on the data without pulse-specific measures of relative amplitude, then on the data including some or all the pulse-specific measures of relative amplitude.
-The model's task was to classifiy each TMS stimulation into one of the four experimental categories.
-In other words, it faced the challenge of establishing to which type of subject, and under which type of session, each stimulation was elicited, based only on the evoked potential, the ISI and ---later--- the relative amplitude measures of the stimulation.
-Our objective was to evaluate whether the inclusion of pulse-specific measures of relative amplitude improved the model's accuracy, and if so in what degree.
-
-The exact line of code that defined the model and its parameters was
-
-\begin{minted}{python}
-    bag_clf = AdaBoostClassifier(
-        DecisionTreeClassifier(max_depth=8), 
-        n_estimators=20, 
-        algorithm='SAMME.R', 
-        learning_rate=0.01)
-\end{minted}
+We trained the exact same random forests model first on the data without
+pulse-specific measures of relative amplitude, then on the data including some
+or all the pulse-specific measures of relative amplitude. The model's task was
+to classifiy each TMS stimulation into one of the four experimental categories.
+In other words, it faced the challenge of establishing to which type of subject,
+and under which type of session, each stimulation was elicited, based only on
+the evoked potential, the ISI and ---later--- the relative amplitude measures of
+the stimulation. Our objective was to evaluate whether the inclusion of
+pulse-specific measures of relative amplitude improved the model's accuracy, and
+if so in what degree.
 
@@
 
-When trained on the raw data, without the inclusion of pulse-specific relative amplitude measures, the model's accuracy was $A\approx34.2 \%$.
+When trained on the raw data, without the inclusion of
+pulse-specific relative amplitude measures, the model's
+accuracy was $A\approx34.2 \%$.
 
+\pagebreak
 \section{Appendix}
@@
 
-\textbf{Theorem: } $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$
+\begin{theorem}
+    $\forall x: x \in \mathbb{R}^+ : \delta(x) \geq \rho(x)$
+\end{theorem}
 
@@
 
-\begin{align} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\ \rho(x) &= \frac{xm}{\sum^m t_j} \end{align}
+\begin{align*} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\
+                \rho(x) &= \frac{xm}{\sum^m t_j} \end{align*}
 
-Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m t_j$. We operate under the assumption that $t_i \in \mathbb{R}^+$. It is the case that
+Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m t_j$. We operate under the
+assumption that $t_i \in \mathbb{R}^+$. It is the case that
 
-\begin{align} \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\
-\iff S_2^mS_1^m &\geq m^2
-\end{align}
+\begin{align*} 
+    \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\ 
+    S_2^mS_1^m &\geq m^2
+\end{align*}
 
-This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So we may assume $S^k_1 S^k_2 \geq k^2$ . We may only show now that
+This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So
+we may assume $S^k_1 S^k_2 \geq k^2$ . We may only show now that
 
-\begin{equation}
+\begin{equation*} 
     S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
-\end{equation}
+\end{equation*}
 
-\begin{align} S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\
+This can be proven as follows.
+
+\begin{align*} 
+    S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\ 
     (S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
@@
     S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k 
-\end{align}
+\end{align*}
 
-We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k + \frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that
+We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k +
+\frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that
 
-\begin{align}\frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
-\iff\frac{1}{t_{k+1}}(t_1+t_2+...+t_k)+t_{k+1}(\frac{1}{t_1}+\frac{1}{t_2}+...+\frac{1}{t_k}) &\geq 2k \\
-\iff \Big(\frac{t_1}{t_{k+1}}+\frac{t_2}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+\frac{t_{k+1}}{t_2}+...+\frac{t_{k+1}}{t_k}\Big) &\geq 2k \\
-\iff \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... + n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
-\end{align}
+\begin{align*}
+    \frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
+    \Big(\frac{t_1}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+...+\frac{t_{k+1}}{t_k}\Big)
+    &\geq 2k \\ \iff
+    \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... +
+    n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
+\end{align*}
 
-We have $\min f=2$  for $f(x)=x+\frac{1}{x}$ for $x \in \mathbb{R}^+$. Then $\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$ for $a,..., n \in \mathbb{R}^+$, which concludes the demonstration.
+We have $\min f=2$  for $f(x)=x+\frac{1}{x}$ for $x \in \mathbb{R}^+$. Then
+$\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$ for $a,..., n \in \mathbb{R}^+$,
+which concludes the demonstration.
 

### 2023-04-14 paper/paper.tex
Paper

\documentclass{article}
\usepackage{amsmath}
\usepackage{ragged2e}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{pmboxdraw}
\usepackage{graphicx} \graphicspath{
{./images/} }
\usepackage{indentfirst}
\usepackage{multicol}
\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
\DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem{wrnote}{Writing note}
\newtheorem{proof}{Proof}

\begin{document}

\begin{multicols}{2}

\section{Introduction}

The emergence of new computational models has provided neuroscientific research
with novel methods for extracting significant insights from experimental results.
This paper takes a computational approach to challenges arising in the analysis
of transcranial stimulation (TMS) data. We argue that conducting TMS experiments
under the double-pulse paradigm enables the use of machine learning models. In
particular, we formulate pulse-specific measures of relative amplitude between
paired and test pulses. These has potential implications in both $a.$ the
production of evidence and $b$. the diagnosis of clinical disorders.

\section{Background}

Transcranial magnetic stimulation (TMS) is a common,
non-invasive experimental technique used to evoke action
potentials in cortical regions of the brain. In particular,
researchers often target the motor cortex and measure the
motor evoked potential (MEP) aroused by the stimulation.
When the purpose of the experiment is to inquire on
neuroplasticity, such stimulations are performed under the
\textit{paired pulse paradigm}.

The \textit{paired pulse paradigm} (or \textit{double pulse
paradigm}) consists in eliciting a series of two temporally
proximate pulses (in the order of miliseconds). The evoked
potentials of the double-stimulation are compared to those
of single test stimulations, and their relative amplitude is
taken as a proxy of neuroplasticity in the brain. The time
separating each of the paired pulses is termed
\textit{interstimulus interval} (ISI). It is the general
case that low intervals (4 or 5 miliseconds) produce
intracortical inhibition, with the evoked potentials of
paired stimulations being generally lower than those of
single pulses. Greater intervals, on the other hand, tend to
produce facilitation.

In the context of this paper, we shall term any coefficient
that serves to represent the proportional relationship
between the potentials evoked by paired and test pulses a
\textit{measure of relative amplitude}.

The goal of this paper is to provide pulse-specific measures of relative
amplitude. This is, coefficients that represent the relative amplitude of each
individual paired pulse with respect to the set of test pulses in an
experimental session. The purpose of this endeavor is to keep data resolution
at its highest, which on its turn allows for the implementation of data-driven
artificial intelligence in the analysis of the 
experimental results. Thus, from a computer science perspective, our objective is
constrained to the sphere of feature engineering.

We will show how pulse-specific measures of relative amplitude allow for
otherwise unfeasible computational analyses of TMS data, such as the use of
machine learning models for the detection of different pulse-response patterns
among different groups of clinical subjects. In particular, we will show they
allow for a machine learning classifier to correctly determine whether a
subject belongs to one of four clinical categories, based only on its evoked
potentials and across different inter-stimulus intervals, with an accuracy of up to
$90\%$.

\section{Data collection}

We used data collected across $N = ?$ subjects at the \textit{Laboratory for the
Study of Slow-wave sleep activity}, University of Pennsylvania, with $H = ?$
healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
Transcranial magnetic stimulation of the motor cortex was elicited to them after
a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
Motor evoked potentials were measured via an electrode (?) placed on the
subjects' thumb (?). In the slow-wave disruption session, an auditory stimulus with
sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet
not strong enough to wake the subjects, was elicited. This experimental setting
produces four distinct categories, two depending on the subject group and two on
the type of sleep session underwent, as shown in the table below.

~

\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
\hline
& Baseline & Slow-wave disruption \\
\hline
Healthy control & HC BL & HC SWD \\
\hline
Major depressive disorder & MDD BL & MDD SWD \\
\hli
\end{tabular}

~

Each observation in the data was a specific experimental sample resulting
from an individual transcranial stimulation. The original features of the data
were:

\begin{enumerate}
    \item An $EMG$ variable with the EMG peak-to-peak of each observation.
    \item A $Label$ categorical feature, which encoded the group of
        the subject of each observation, was the target variable. 
    \item An $ISI$ feature that econded the inter-stimulus interval of each
        pulse. A value of $-1$ indicated that the given pulse was a test pulse (no
        inter-stimulus interval).
\end{enumerate}

\section{Limitations of the traditional analysis}

Measures of relative amplitude in neuroscience are generally computed at the
subject level by repeatedly taking averages in the following manner. First, the
quotient of the average paired-pulse and the average test pulse is found per
every subject. Secondly, those quotients are averaged out across all subjects in
a subject group. This is reasonable, since hypotheses generally deal with
differences across subject groups.

Two major limitations arise from this method. The principle one is the fact that
it implies greatly downscaling the data resolution Secondly, potentials evoked
by test and paired pulses present a very high variance (Rossini et. al., 20015;
Orth, Snijders and Rothwell, 2003; Wassermann, 2002), which —being that it is a
quotient of two means—  raises questions about the quality of the representation
where relatively few pulses are elicited.

When considering the size of the data across multiple experimental subjects,
this shrinkage in data resolution is far from being negligible. And in times
when scientific inquiry can greatly benefit from forms of artificial
intelligence that rely on large amounts of data (e.g. machine learning), it is
clearly limiting to only count with low-resolution measures to interpret
experimental results.

\section{Proposed approach}

\subsection{Pulse-specific relative amplitude measures}

Considering the limitations of the traditional approach, this paper proposes two
pulse-specific relative amplitude measures. This allows for otherwise unfeasible
computational analyses of TMS data, such as the use of machine learning (ML) models
for the detection of different pulse-response patterns among different groups of
clinical subjects.

To empirically test the validity of these measures, we evaluated whether they
improved the performance of a classification model, and in what degree. (We
chose a classification model for practical reasons --namely, because the
samples in our data were conveniently labeled by subject group and sleep
session).

In order to do this, we defined a random forest model in Python under the
following parameters: \ldots

The model was tasked with classifying every observation with the appropriate subject
label. In other words, it was set out to infer, based on the properties of each
transcranial stimulation response, the group of the subject upon which the
stimulation was elicited, as well as the type of sleep session after which it
occurred.

The model was first trained on the raw data, with the features described in
\textbf{3.1}. Then, it was trained with the inclusion of one or many engineered
features of relative amplitude (see section \textbf{4}). These features were
computed for each observation using the Julia programming language.

Since the neural response to transcranial stimulations follows an exponential
distribution (see \textbf{Cite section}), we experimented with the inclusion of
the features above applied to the logarithmically transformed peak-to-peak EMG.

As stated earlier, action potentials evoked by transcranial magnetic
stimulations follow a Gamma distribution that is very close to the exponential.
(see \textbf{Statistics} section). Experimentation has shown that the inclusion
of weighted variants of the features defined above greatly improves the
performance of a random forests model (see \textbf{Empirical results}).

The use of weighted instead of arithmetic averages may be useful in dealing with
the excessive influence of outliers or highly spread out points in the feature.
For example, the weight vectors may be computed using the MAD or the
inverse-variance of each $x$.

Thus, our hypothesis is that pulse-specific measures of relative
amplitude will significantly improve a random forest's performance.

The following section (\textbf{Cite section}) defines the proposed measures of
relative amplitude. In section \textbf{Results}, we show the impact of these
measures on the performance of the random forest model.

\subsection{Definitions: Pulse specific relative amplitudes}

For simplicity, we will deal with the hypothetical situation
in which a single ISI was used for paired stimulations. All
of our results generalize to different inter-stimulus intervals.

Let $k$ be the number of experimental subjects in some
subject group $\mathcal{G}$, to each of whom $n$ paired stimulations and
$m$ test stimulations were elicited.

\begin{definition}

Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices over the
vector space $\mathbb{R}^+$ representing
the paired and test potentials evoked across each of the $k$ subjects.

\end{definition}

\begin{definition} 
    Let $x \in \textbf{P}_{\star i}$ be the MEP of a single paired stimulation
    elicited on the $i$th experimental subject, and $\textbf{t} =
    \textbf{T}_{\star i}$ the vector containing the MEPs of all test pulses
    elicited on that subject. Let $\textbf{w}$ be some appropriate weight
    vector. Then we define two pulse-specific relative amplitude measures,

    \begin{equation} 
        \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
    \end{equation}

    \begin{equation} 
        \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} 
    \end{equation}

    \begin{equation}
            \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\ 
    \end{equation}

    \begin{equation}
        \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^mw_j} 
    \end{equation}
\end{definition}

\begin{remark} 
    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
    property, consult the appendix.) 
\end{remark}

Notice that $\rho(x)$ is the proportion between the potential $x$, evoked by a
paired stimulation, with respect to the average potential of single test
stimulations. On the other hand, $\delta(x)$ is the average proportion of $x$
with respect to each single test pulse.

Notice as well that $\rho$ and $\delta$ are defined over the test pulses in a
specific experimental session $i$. In other words, whenever we speak of the
$\delta$ and $\rho$ \textit{features} (\textit{id est}, of the data resulting from
broadcasting $\rho$ and $\delta$ over a series of observations), such feature is
understood to be subject- and session-specific.

In particular, $\rho$ is a representation of the relative importance of each
double pulse in relation to the overall distribution of the test pulses in the
subject. It measures the deviation of the double pulse $x$ with respect to the
average test pulse. On the other hand, $\delta$ is a measure of the
proportionality of a double pulse with respect to different values in the
distribution of the test pulses of the subject.

\section{Results}

\subsection{ML model without relative amplitude measures}

When trained on the raw data, without the inclusion of pulse-specific relative
amplitude measures, the model's accuracy was of $\approx34.2 \%$. However, the
confusion matrix of the model shows the errors are concentrated on the healthy
control categories, while categorization of diagnosed subjects was substantially
better. This implies major depressive subjects show statistically significant
and distinct patterns in their TMS responses in comparison to healthy controls,
and may be considered evidence for depression-induced differences in
neuroplasticity.

\end{multicols}

\centering
~

\includegraphics[scale=0.4]{cm-raw}

\justifying
\begin{multicols}{2}

However, it is important to note that the proxy for neuroplasticity in the
double pulse paradigm is precisely the relative amplitude of double pulses with
respect to test pulses. The raw data lacks any measure of relative amplitude.
Besides, regardless of the fact that the model above suggests a critical
difference in the response patterns between subject groups, its accuracy is
still poor.

\subsection{Engineered data}

With the inclusion of the $\rho$ feature, accuracy increases by a factor of
$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature alone increased
it, in comparison to the raw data, had a more or less equivalent impact,
increasing accuracy to $73.4\%$.

\end{multicols}
\raggedright

~

\includegraphics[scale=0.24]{cm-rho}
\raggedright
\includegraphics[scale=0.24]{cm-delta}

\justifying
\begin{multicols}{2}

The model still shows a higher accuracy when classifying diagnosed subjects, but
the overall accuracy across all categories was significantly improved.

As stated earlier, the inclusion of the $\delta$ and $\rho$ features computed
over the logarithmically transformed EMG peak-to-peak improved the model's
accuracy. Concretely, accuracy was increased to $86\%$, with the following
confusion matrix.

\end{multicols}

~
\centering

\includegraphics[scale=0.4]{cm-dr-ln}

\justifying
\begin{multicols}{2}

At last, if to the previous model we add also the weighted versions of $\delta$
and $\rho$, using inverse-variance weights, we obtained an accuracy of $89.8\%$,
with the following confusion matrix.

\end{multicols}

\centering
~

\includegraphics[scale=0.4]{final-cm}

\justifying
\begin{multicols}{2}

\section{Statistics}

\begin{wrnote}
    The description of the experimental process in this section is poor. Also I
    do not know the total number of subjects, experiments are still being
    conducted. Need input from the lab on these matters.
\end{wrnote}

Although distributions were always exponential, the $\beta$ parameter of said
distributions varied across subject groups and session types.

~

    \begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
    \hline
    &       Mean & Median \\
    \hline
    HC BL & 196.77 & 149.36 \\
    \hline
    HC SWD & 165.98 & 99.76 \\
    \hline \\ 
    MDD BL & 187.56 & 127.69\\ 
    \hline 
    MDD SWD & 247.57 & 151.07 \\
    \hline
    \end{tabular}

~

Each of the means in the above table correspond to the estimated $\beta$
parameter of the distribution of the paired pulses on each subject group.

\end{multicols}
\raggedright

~

\includegraphics[scale=0.14]{Paired-TestDist}
\raggedright
\includegraphics[scale=0.14]{P-dist-subjects-big}

\justifying
\begin{multicols}{2}

\section{Discussion}

The previous results show that the inclusion of pulse-specific relative
amplitude measures greatly improve the accuracy of a random forests model
trained over TMS data for subject classification. More importantly, it becomes
clear that, via the engineered features $\delta$ and $\rho$, researchers can
attain evidence in favour or against hypotheses pertaining to differences among
subject groups. Particularly, in showing that pulse-specific measures of
relative amplitude are significant information to determine the sleep session
and group of a subject, and being relative amplitude measures a proxy for
neuroplasticity, the previous model provides evidence in favour of
sleep-modulated, depression-induced differences in neuroplasticity.

The use of machine learning models over neuroscientific observations is not only
a promising tool in the production of evidence. There is a diagnostic potential
that is still to be evaluated. Indeed, if machine learning models can detect the
neural patterns that distinguish, under specific experimental conditions,
healthy control from diagnosed subjects, such models can potentially be
implemented in the diagnosis process as powerful clinical tool.

Although long and serious scientific effort is still required to appraise the
diagnostic potential of machine learning models, we believe our results allow
for a certain amount of conservative optimism on the matter.

In short, pulse-specific relative amplitude features are successful in making
machine learning models applicable to TMS experimental data. Thus, they can play
an important part in future research by allowing for new ways of analyzing and
extracting meaningful information of TMS results.

\section{Appendix 1}

\section{Appendix 2}

\textbf{Proof 1}. In \textbf{Remark 1}, we observed the following property:

    \begin{align*}
        \forall x: x \in \mathbb{R}^+ : \delta(x) \geq \rho(x)
    .\end{align*}

Such property can be proven via induction. Firstly, recall that

\begin{align*} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\
                \rho(x) &= \frac{xm}{\sum^m t_j} \end{align*}

Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m t_j$. We operate under the
assumption that $t_i \in \mathbb{R}^+$. It is the case that

\begin{align*} 
    \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\ 
    S_2^mS_1^m &\geq m^2
\end{align*}

This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So
we may assume $S^k_1 S^k_2 \geq k^2$ . We now set out to show that

\begin{equation*} 
    S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
\end{equation*}

This can be proven as follows.

\begin{align*} 
    S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\ 
    (S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
    S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2+1 &\geq k^2+2k+1 \\
    S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k 
\end{align*}

We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k +
\frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that

\begin{align*}
    \frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
    \Big(\frac{t_1}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+...+\frac{t_{k+1}}{t_k}\Big)
    &\geq 2k \\ \iff
    \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... +
    n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
\end{align*}

We have $\min f=2$  for $f(x)=x+\frac{1}{x}$ for $x \in \mathbb{R}^+$. Then
$\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$ for $a,..., n \in \mathbb{R}^+$,
which concludes the demonstration.

\end{document}

### 2023-04-14 research/notes.tex
Paper

@@
 \usepackage{pmboxdraw}
-\usepackage{minted}
-\usepackage{graphicx}
-\graphicspath{ {./images/} }
+\usepackage{graphicx} \graphicspath{
+{./images/} }
+\usepackage{indentfirst}
+\usepackage{multicol}
+\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
 \DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}
@@
 \begin{document}
+
+\begin{multicols}{2}
 \section{Introduction}
@@
 
-Transcranial magnetic stimulation (TMS) is a common, non-invasive experimental technique used to evoke action potentials in cortical regions of the brain.
-In particular, researchers often target the motor cortex and measure the motor evoked potential (MEP) aroused by the stimulation.
-When the purpose of the experiment is to inquire on neuroplasticity, such stimulations are performed under a \textit{paired-pulse paradigm}.
+Transcranial magnetic stimulation (TMS) is a common,
+non-invasive experimental technique used to evoke action
+potentials in cortical regions of the brain. In particular,
+researchers often target the motor cortex and measure the
+motor evoked potential (MEP) aroused by the stimulation.
+When the purpose of the experiment is to inquire on
+neuroplasticity, such stimulations are performed under a
+\textit{paired-pulse paradigm}.
 
-The \textit{paired-pulse paradigm} (or \textit{double pulse paradigm}) consists in eliciting a series of two temporally proximate pulses (in the order of miliseconds).
-The evoked potentials of the double-stimulations are compared to those of single test stimulations, and their relative amplitude is taken as a proxy of neuroplasticity in the brain.
-The time separating each of the paired-pulses is termed \textit{interstimulous interval} (ISI).
-It is the general case that low intervals (4 or 5 miliseconds) produce intracortical inhibition, with the evoked potentials of paired stimulations being generally lower than those of single pulses.
-Greater intervals, on the other hand, tend to produce facilitation.
+The \textit{paired-pulse paradigm} (or \textit{double pulse
+paradigm}) consists in eliciting a series of two temporally
+proximate pulses (in the order of miliseconds). The evoked
+potentials of the double-stimulations are compared to those
+of single test stimulations, and their relative amplitude is
+taken as a proxy of neuroplasticity in the brain. The time
+separating each of the paired-pulses is termed
+\textit{interstimulous interval} (ISI). It is the general
+case that low intervals (4 or 5 miliseconds) produce
+intracortical inhibition, with the evoked potentials of
+paired stimulations being generally lower than those of
+single pulses. Greater intervals, on the other hand, tend to
+produce facilitation.
 
-In the context of this paper, we shall term any coefficient that serves to represent the proportional relationship between the potentials evoked by paired and test pulses a \textit{measure of relative amplitude}.
-Measures of relative amplitude in neuroscience are generally computed at the subject level.
-This is reasonable, since hypotheses generally deal with differences across subject groups.
-However, TMS experimental results are pulse-specific evoked potentials, and transforming them to subject- or group-specific measures implies lowering the data resolution (i.e.
-representing the same phenomena with less information).
-In times when scientific inquiry can greatly benefit from forms of artificial intelligence that rely on large amounts of data, it is limiting to count only with low-resolution measures.
+In the context of this paper, we shall term any coefficient
+that serves to represent the proportional relationship
+between the potentials evoked by paired and test pulses a
+\textit{measure of relative amplitude}. Measures of relative
+amplitude in neuroscience are generally computed at the
+subject level. This is reasonable, since hypotheses
+generally deal with differences across subject groups.
+However, TMS experimental results are pulse-specific evoked
+potentials, and transforming them to subject- or
+group-specific measures implies down-scaling the data resolution.
 
-Hence, the goal of this paper is to provide pulse-specific measures of relative amplitude.
-We formulate coefficients that represent the relative amplitude of individual paired-pulses, keeping data resolution at its highest.
-Thus, from a computational perspective, our objective is constrained to the sphere of feature engineering.
-This allows for otherwise unfeasible computational analyses of TMS data, such as the use of machine learning models for the detection of different pulse-response patterns among different groups of clinical subjects.
+The goal of this paper is to provide pulse-specific measures of relative
+amplitude. This is, coefficients that represent the relative amplitude of each
+individual paired-pulse with respect to the set of test pulses in an
+experimental session. The purpose of this endeavor is to keep data resolution
+at its highest, which on its turn allows for the implementation of data-driven
+artificial intelligence in the analysis of the 
+experimental results. Thus, from a computational perspective, our objective is
+constrained to the sphere of feature engineering.
 
-We show such features allow for a machine learning classifier to correctly determine whether a subject belongs to one of four clinical categories based only on its evoked potentials across different inter-stimulous intervals with an accuracy of up to $90\%$.
+We shall show how pulse-specific measures of relative amplitude allow for
+otherwise unfeasible computational analyses of TMS data, such as the use of
+machine learning models for the detection of different pulse-response patterns
+among different groups of clinical subjects. In particular, we will show they
+allow for a machine learning classifier to correctly determine whether a
+subject belongs to one of four clinical categories based only on its evoked
+potentials across different inter-stimulous intervals with an accuracy of up to
+$90\%$.
 
@@
 
-For simplicity, we will deal with the hypothetical situation in which a single ISI was used for paired stimulations. All of our results generalize to different ISI.
+For simplicity, we will deal with the hypothetical situation
+in which a single ISI was used for paired stimulations. All
+of our results generalize to different ISI.
 
@@
 
-Let $k$ be the number of experimental subjects in some subject group, to each of whom $n$ paired stimulations and $m$ test stimulations were elicited.
+Let $k$ be the number of experimental subjects in some
+subject group, to each of whom $n$ paired stimulations and
+$m$ test stimulations were elicited.
 
@@
 
-    Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices representing the paired and test potentials evoked across each of the $k$ subjects, such that
+    Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be
+    matrices representing the paired and test potentials
+    evoked across each of the $k$ subjects, such that
 
-    \begin{equation}
-    \textbf{P} := \begin{bmatrix}x_{11} & x_{12}& \cdots & x_{1k} \\
-    x_{21} & x_{22}& \cdots & x_{2k} \\
-    & & ... & \\
-    x_{n1} & x_{n2}& \cdots & x_{nk} 
-    \end{bmatrix} ~~
-    \textbf{T}:= \begin{bmatrix}t_{11} & t_{12}& ... & t_{1k} \\
-    t_{21} & t_{22}& \cdots & t_{2k} \\
-    & & ... & \\
-    t_{m1} & t_{m2}& \cdots & t_{mk} 
-    \end{bmatrix}
-    \end{equation}
-\end{definition}
+    \begin{equation} \textbf{P} := \begin{bmatrix}x_{11} &
+    x_{12}& \cdots & x_{1k} \\ x_{21} & x_{22}& \cdots &
+x_{2k} \\ & & ... & \\ x_{n1} & x_{n2}& \cdots & x_{nk}
+\end{bmatrix} ~~ \textbf{T}:= \begin{bmatrix}t_{11} &
+t_{12}& ... & t_{1k} \\ t_{21} & t_{22}& \cdots & t_{2k} \\
+      & & ... & \\ t_{m1} & t_{m2}& \cdots & t_{mk}
+\end{bmatrix} \end{equation} \end{definition}
 
-\begin{definition}
-Let $\mu_t := \begin{bmatrix} \mu_1, \mu_2, \ldots, \mu_k \end{bmatrix}^\top$ be a vector where $\mu_i$ is the average test potential for the $i$th subject. In other words,
+        \begin{definition} Let $\mu_t := \begin{bmatrix} \mu_1 &
+        \mu_2 & \ldots & \mu_k \end{bmatrix}^\top$ be such that 
+$\mu_i$ is the average test potential for the $i$th subject.
+In other words,
 
-\begin{equation}
-    \mu_i = \frac{1}{m}\sum_{j=1}^m t_{ji}
-\end{equation}
-\end{definition}
+\begin{equation} \mu_i = \frac{1}{m}\sum_{j=1}^m t_{ji}
+\end{equation} \end{definition}
 
-\begin{remark}
-    Evoked potentials of both paired and test pulses conform to a Gamma distribution, as discussed in the appendix (?).
-This means it is always assumed $\forall x \in \textbf{P}, \forall t\in \textbf{T}|x, t \in \mathbb{R}^+$.
-\end{remark}
+\begin{remark} Evoked potentials of both paired and test
+    pulses conform to a Gamma distribution, as discussed in
+    the appendix (?). This means it is always an implicit assumption that
+    $\forall (x \in \textbf{P}, t\in \textbf{T})|x, t
+\in \mathbb{R}^+$. \end{remark}
 
@@
 
-\begin{definition}
-Let $x \in \textbf{P}_{\star i}$ be a single paired-stimulation MEP in the $i$th experimental subject, and $\textbf{t} = \textbf{T}_{\star i}$ the vector containing all test MEPs of that subject.
-Then we define two pulse-specific relative amplitude measures,
+\begin{definition} Let $x \in \textbf{P}_{\star i}$ be a
+    single paired-stimulation MEP in the $i$th experimental
+    subject, and $\textbf{t} = \textbf{T}_{\star i}$ the
+    vector containing all test MEPs of that subject. Then we
+    define two pulse-specific relative amplitude measures,
 
-\begin{equation}
-    \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
+    \begin{equation} \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
     \end{equation}
 
-\begin{equation}
-    \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j}
-\end{equation}
+    \begin{equation} \delta(x) :=
+    \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} \end{equation}
 \end{definition}
 
-\begin{remark}
-    $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$. (For a proof of this property, consult the appendix.)
-\end{remark}
+\begin{remark} $\forall x \in \mathbb{R}^+|\delta(x) \geq
+\rho(x)$. (For a proof of this property, consult the
+appendix.) \end{remark}
 
-Notice that $\rho$ is the proportion between the potential $x$, evoked by a paired stimulation, with respect to the average potential of single test stimulations.
-On the other hand, $\delta$ is the average proportion of $x$ with respect to each single test pulse.
+Notice that $\rho$ is the proportion between the potential
+$x$, evoked by a paired stimulation, with respect to the
+average potential of single test stimulations. On the other
+hand, $\delta$ is the average proportion of $x$ with respect
+to each single test pulse.
 
-The $\rho$ function is not an entirely new contribution.
-The traditional group level relative amplitude measure is conceived as the average, across all subjects in a group, of the average paired response divided by the average test response of each subject.
-If we let $S_i$ be the average relative amplitude of the $i$th subject, then $S_i$ has ben traditionally defined as
+The $\rho$ function is not an entirely new contribution. The
+traditional group level relative amplitude measure is
+conceived as the average, across all subjects in a group, of
+the average paired response divided by the average test
+response of each subject. If we let $S_i$ be the average
+relative amplitude of the $i$th subject, then $S_i$ has ben
+traditionally defined as
 
-\begin{equation}
-    S_i=\frac{\frac{1}{n}\sum_{j=1}^n x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
-\end{equation}
+\begin{equation} S_i=\frac{\frac{1}{n}\sum_{j=1}^n
+x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} =
+\frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n} \end{equation}
 
-as it is easy to see from decomposing the sum in the numerator.
-In other words, the traditionally used measure of relative amplitude, at the subject level, has always been the average $\rho$ in a subject.
+as it is easy to see from decomposing the sum in the
+numerator. In other words, the traditionally used measure of
+relative amplitude, at the subject level, has always been
+the average $\rho$ in a subject.
 
@@
 
-As stated earlier, action potentials evoked by trasncranial magnetic stimulation follow a Gamma distribution; one that closely resembles an exponential distribution (see \textbf{Statistics} section).
-A valid form of data augmentation that also attempts to deal with the presence of outliers is to weight the averages involved in $\rho, \delta$ using inverse-variance weights.
-The attempt is to allow for potentials that lay at the tail of the distribution to contribute in a way proportional to their probability.
+As stated earlier, action potentials evoked by trasncranial
+magnetic stimulation follow a Gamma distribution; one that
+closely resembles an exponential distribution (see
+\textbf{Statistics} section). A valid form of data
+augmentation that also attempts to deal with the presence of
+outliers is to weight the averages involved in $\rho,
+\delta$ using inverse-variance weights. The attempt is to
+allow for potentials that lay at the tail of the
+distribution to contribute in a way proportional to their
+probability.
 
-If we let $\rho_w, \delta_w$ be the weighted versions of $\rho, \delta$, then we have
+If we let $\rho_w, \delta_w$ be the weighted versions of
+$\rho, \delta$, then we have
 
-\begin{align}
-    \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\
-    \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m w_j}
-\end{align}
+        \begin{align} \rho_w(x) &:= \frac{xm\sum_{j=1}^m
+        w_j}{\sum_{j=1}^m t_j w_j} \\ \delta_w(x) &:=
+    \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m
+w_j} \end{align}
 
-where $\textbf{w}$ is some appropriate weight vector.
-Since the purpose of $\textbf{w}$ is to reduce the contribution of outliers to the overall measure of relative amplitude, it is to be expected that the distribution of $\textbf{w}$ closely resembles that of evoked potentials.
+where $\textbf{w}$ is some appropriate weight vector. Since
+the purpose of $\textbf{w}$ is to reduce the contribution of
+outliers to the overall measure of relative amplitude, it is
+to be expected that the distribution of $\textbf{w}$ closely
+resembles that of evoked potentials.
 
@@
 
-We used data collected across $N = ?$ subjects at the \textit{Laboratory for the Study of Slow-wave sleep}, University of Pennsylvania, with $H = ?$ healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
-Transcranial magnetic stimulation of the motor cortex was elicited to them after a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
-Motor evoked potentials were measured via an electrode (?) placed on the subjects' thumb.
-In the slow-wave disruption session, an auditory stimulus with sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet not strong enough to wake the subjects, was elicited.
-This experimental setting produces four distinct categories, two depending on the subject group and two on the type of sleep session underwent.
+We used data collected across $N = ?$ subjects at the
+\textit{Laboratory for the Study of Slow-wave sleep activity},
+University of Pennsylvania, with $H = ?$ healthy controls
+and $D = ?$ diagnosed with major depressive disorder (MDD).
+Transcranial magnetic stimulation of the motor cortex was
+elicited to them after a night of baseline sleep and after a
+night of slow-wave disruption (SWD) sleep. Motor evoked
+potentials were measured via an electrode (?) placed on the
+subjects' thumb. In the slow-wave disruption session, an
+auditory stimulus with sufficient strength to interrupt the
+normal occurrence of slow-wave sleep, yet not strong enough
+to wake the subjects, was elicited. This experimental
+setting produces four distinct categories, two depending on
+the subject group and two on the type of sleep session
+underwent.
 
-We trained the exact same random forests model first on the data without pulse-specific measures of relative amplitude, then on the data including some or all the pulse-specific measures of relative amplitude.
-The model's task was to classifiy each TMS stimulation into one of the four experimental categories.
-In other words, it faced the challenge of establishing to which type of subject, and under which type of session, each stimulation was elicited, based only on the evoked potential, the ISI and ---later--- the relative amplitude measures of the stimulation.
-Our objective was to evaluate whether the inclusion of pulse-specific measures of relative amplitude improved the model's accuracy, and if so in what degree.
-
-The exact line of code that defined the model and its parameters was
-
-\begin{minted}{python}
-    bag_clf = AdaBoostClassifier(
-        DecisionTreeClassifier(max_depth=8), 
-        n_estimators=20, 
-        algorithm='SAMME.R', 
-        learning_rate=0.01)
-\end{minted}
+We trained the exact same random forests model first on the
+data without pulse-specific measures of relative amplitude,
+then on the data including some or all the pulse-specific
+measures of relative amplitude. The model's task was to
+classifiy each TMS stimulation into one of the four
+experimental categories. In other words, it faced the
+challenge of establishing to which type of subject, and
+under which type of session, each stimulation was elicited,
+based only on the evoked potential, the ISI and ---later---
+the relative amplitude measures of the stimulation. Our
+objective was to evaluate whether the inclusion of
+pulse-specific measures of relative amplitude improved the
+model's accuracy, and if so in what degree.
 
@@
 
-\textbf{Theorem: } $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$
+\textbf{Theorem: } $\forall x \in \mathbb{R}^+|\delta(x)
+\geq \rho(x)$
 
@@
 
-\begin{align} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\ \rho(x) &= \frac{xm}{\sum^m t_j} \end{align}
+                \begin{align} \delta(x) &=
+                \frac{x}{m}\sum^m\frac{1}{t_j} \\ \rho(x) &=
+            \frac{xm}{\sum^m t_j} \end{align}
 
-Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m t_j$. We operate under the assumption that $t_i \in \mathbb{R}^+$. It is the case that
+            Let $S_1^m = \sum^m\frac{1}{t_j}, S_2^m= \sum^m
+            t_j$. We operate under the assumption that $t_i
+            \in \mathbb{R}^+$. It is the case that
 
-\begin{align} \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\
-\iff S_2^mS_1^m &\geq m^2
+                    \begin{align}
+                    \frac{x}{m}\sum^m\frac{1}{t_j} &\geq
+                \frac{xm}{\sum^m t_j} \\ \iff S_2^mS_1^m
+                                                   &\geq m^2
                 \end{align}
 
-This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So we may assume $S^k_1 S^k_2 \geq k^2$ . We may only show now that
+                This holds for $m=1$, since
+                $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq
+                t_1$. So we may assume $S^k_1 S^k_2 \geq
+                k^2$ . We may only show now that
 
-\begin{equation}
-S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
-\end{equation}
+                \begin{equation} S^{k+1}_1 S^{k+1}_2 \geq
+                (k+1)^2 \end{equation}
 
-\begin{align} S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\
-(S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
-S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2+1 &\geq k^2+2k+1 \\
-S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k
+                    \begin{align} S^{k+1}_1 S^{k+1}_2 &\geq
+                        (k+1)^2 \\
+                        (S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1})
+                        &\geq k^2+2k+1 \\ S^k_1S^k_2+
+                        t_{k+1}S_1^k +
+                        \frac{1}{t_{k+1}}S^k_2+1 &\geq
+                        k^2+2k+1 \\ S^k_1S^k_2+ t_{k+1}S_1^k
+                    + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k
                 \end{align}
 
-We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k + \frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that
+                We know $S^k_1S^k_2 \geq k^2$ and then it
+                suffices to show $t_{k+1}S_1^k +
+                \frac{S^k_2}{t_{k+1}}\geq 2k$. To prove
+                this, simply observe that
 
-\begin{align}\frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
-\iff\frac{1}{t_{k+1}}(t_1+t_2+...+t_k)+t_{k+1}(\frac{1}{t_1}+\frac{1}{t_2}+...+\frac{1}{t_k}) &\geq 2k \\
-\iff \Big(\frac{t_1}{t_{k+1}}+\frac{t_2}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+\frac{t_{k+1}}{t_2}+...+\frac{t_{k+1}}{t_k}\Big) &\geq 2k \\
-\iff \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... + n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
-\end{align}
+                        \begin{align}\frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}}
+                        &\geq 2k \\
+                        \iff\frac{1}{t_{k+1}}(t_1+t_2+...+t_k)+t_{k+1}(\frac{1}{t_1}+\frac{1}{t_2}+...+\frac{1}{t_k})
+                        &\geq 2k \\ \iff
+                        \Big(\frac{t_1}{t_{k+1}}+\frac{t_2}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+\frac{t_{k+1}}{t_2}+...+\frac{t_{k+1}}{t_k}\Big)
+                        &\geq 2k \\ \iff
+                        \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+...
+                    + n+\frac{1}{n}}^{\text{$2k$ terms} }
+                        &\geq 2k \end{align}
 
-We have $\min f=2$  for $f(x)=x+\frac{1}{x}$ for $x \in \mathbb{R}^+$. Then $\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$ for $a,..., n \in \mathbb{R}^+$, which concludes the demonstration.
+                        We have $\min f=2$  for
+                        $f(x)=x+\frac{1}{x}$ for $x \in
+                        \mathbb{R}^+$. Then
+                        $\min(a+\frac{1}{a}+...+n+\frac{1}{n})=2k$
+                        for $a,..., n \in \mathbb{R}^+$,
+                        which concludes the demonstration.
 

### 2023-04-14 research/v2.tex
Paper

@@
 \usepackage{amsmath}
+\usepackage{ragged2e}
 \usepackage{amssymb}
@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{wrnote}{Writing note}
+\newtheorem{proof}{Proof}
 
@@
 
-\begin{definition} 
-    Let $\mu_t := \begin{bmatrix} \mu_1 & \mu_2 & \ldots & \mu_k
-    \end{bmatrix}^\top$ be such that $\mu_i$ is the average test potential for
-    the $i$th subject. In other words,
-
-\begin{equation} 
-    \mu_i = \frac{1}{m}\sum_{j=1}^m t_{ji}
-\end{equation} \end{definition}
-
-\begin{remark} 
-    Evoked potentials of both paired and test pulses conform to a Gamma
-    distribution, as discussed in the appendix (?). This means it is always an
-    implicit assumption that $\langle\forall x, t : x \in \textbf{P}, t\in \textbf{T}
-    : x, t \in \mathbb{R}^+\rangle$. 
-\end{remark}
-
 \subsection{The $\rho$ and $\delta$ features}
@@
 
-The $\rho$ function is not an entirely new contribution. Traditionally, the
-group-level relative amplitude measure is conceived as the average, across all
-subjects in a group, of the average paired response divided by the average test
-response. This means that, if we let $S_i$ be the average relative amplitude of
-the $i$th subject, then $S_i$ has been traditionally defined as
+Each feature improves the performance of a random forests model in a similar
+degree, as will be shown later. This corroborates that both capture different
+but complementary information ---as it is intended in their formulations. In
+particular, $\rho$ can be a useful representation of the relative importance of
+each double pulse in relation to the overall distribution of the test pulses in
+the subject. It measures the deviation of the double pulse $x$ with respect to
+the average test pulse. On the other hand, $\delta$ is a measure of the
+proportionality of a double pulse with respect to different values in the
+(certainly exponential) distribution of the test pulses in the subject.
 
-\begin{equation} 
+Since the neural response to transcranial stimulations follows a distribution
+$\beta$ very close to exponential (see \textbf{Statistics} section), we
+experimented with the inclusion of the features above applied to the
+logarithmically transformed MEPs. The inclusion of both $\rho(x), \rho(\ln(x))$
+for every $x$ improves the performance of a random forest model significantly,
+and the same occurs for $\delta$, as can be seen in the \textbf{Empirical
+results} section.
+
+It must be said that the $\rho$ function is not an entirely new contribution.
+Traditionally, the group-level relative amplitude measure is conceived as the
+average, across all subjects in a group, of the average paired response divided
+by the average test response. This means that, if we let $S_i$ be the average
+relative amplitude of the $i$th subject, then $S_i$ has been traditionally
+defined as
+
+\begin{equation*} 
     S_i=\frac{\frac{1}{n}\sum_{j=1}^n x_{ji}}{\frac{1}{m}\sum_{j=1}^m t_{ji}} = 
     \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n} 
-\end{equation}
+\end{equation*}
 
@@
 
-As stated earlier, action potentials evoked by transcranial
-magnetic stimulations follow a Gamma distribution; one that
-closely resembles an exponential distribution (see
-\textbf{Statistics} section). A valid form of data
-augmentation that also attempts to deal with the presence of
-outliers is to weight the averages involved in $\rho,
-\delta$ using inverse-variance weights. The attempt is to
-allow for potentials that lay at the tail of the
-distribution to contribute in a way proportional to their
-probability.
+As stated earlier, action potentials evoked by transcranial magnetic
+stimulations follow a Gamma distribution that is very close to the exponential.
+(see \textbf{Statistics} section). Experimentation has shown that the inclusion
+of weighted variants of the features defined above greatly improves the
+performance of a random forests model (see \textbf{Empirical results}).
 
-If we let $\rho_w, \delta_w$ be the weighted versions of
-$\rho, \delta$, then we have
+The use of weighted instead of arithmetic averages may be useful in dealing with
+the excessive influence of outliers or highly spread out points in the feature.
+For example, the weight vectors may be computed using the MAD or the
+inverse-variance of each $x$.
 
-        \begin{align} \rho_w(x) &:= \frac{xm\sum_{j=1}^m
-        w_j}{\sum_{j=1}^m t_j w_j} \\ \delta_w(x) &:=
-    \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^m
-w_j} \end{align}
+In our \textbf{Empirical results}, we show the influence of including
+inverse-variance variance weights to the model. But the general formulation of
+these alternative features, for any desired weight vector, is
 
-where $\textbf{w}$ is some appropriate weight vector. Since
-the purpose of $\textbf{w}$ is to reduce the contribution of
-outliers to the overall measure of relative amplitude, it is
-to be expected that the distribution of $\textbf{w}$ closely
-resembles that of evoked potentials.
+\begin{align} 
+    \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\ 
+    \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^mw_j} 
+\end{align}
 
-\section{Empirical results}
+where $\textbf{w}$ is some appropriate weight vector.
+
+\section{Statistics}
+
+\begin{wrnote}
+    The description of the experimental process in this section is poor. Also I
+    do not know the total number of subjects, experiments are still being
+    conducted. Need input from the lab on these matters.
+\end{wrnote}
 
@@
 Motor evoked potentials were measured via an electrode (?) placed on the
-subjects' thumb. In the slow-wave disruption session, an auditory stimulus with
+subjects' thumb (?). In the slow-wave disruption session, an auditory stimulus with
 sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet
@@
 produces four distinct categories, two depending on the subject group and two on
-the type of sleep session underwent.
+the type of sleep session underwent, as shown in the table below.
 
-We trained the exact same random forests model first on the data without
-pulse-specific measures of relative amplitude, then on the data including some
-or all the pulse-specific measures of relative amplitude. The model's task was
-to classifiy each TMS stimulation into one of the four experimental categories.
-In other words, it faced the challenge of establishing to which type of subject,
-and under which type of session, each stimulation was elicited, based only on
-the evoked potential, the ISI and ---later--- the relative amplitude measures of
-the stimulation. Our objective was to evaluate whether the inclusion of
-pulse-specific measures of relative amplitude improved the model's accuracy, and
-if so in what degree.
+~
+
+\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
+\hline
+& Baseline & Slow-wave disruption \\
+\hline
+Healthy control & HC BL & HC SWD \\
+\hline
+Major depressive disorder & MDD BL & MDD SWD \\
+\hline
+\end{tabular}
+
+~
+
+In our statistical analyses, we have contemplated test and paired pulses
+separately, since they are different type of stimulations. We have found the
+peak-to-peak EMG to follow an exponential distribution in both test and paired
+pulses. This is true when taking different inter-stimulus intervals in
+consideration as well as when observing the distribution of distinctly spaced
+paired pulses.
+
+Although distributions were always exponential, the $\beta$ parameter of said
+distributions varied across subject groups and session types.
+
+~
+
+    \begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|  }
+    \hline
+    &       Mean & Median \\
+    \hline
+    HC BL & 196.77 & 149.36 \\
+    \hline
+    HC SWD & 165.98 & 99.76 \\
+    \hline \\ 
+    MDD BL & 187.56 & 127.69\\ 
+    \hline 
+    MDD SWD & 247.57 & 151.07 \\
+    \hline
+    \end{tabular}
+
+~
+
+Each of the means in the above table correspond to the estimated $\beta$
+parameter of the distribution of the paired pulses on each subject group.
+
+\end{multicols}
+\raggedright
+
+~
+
+\includegraphics[scale=0.14]{Paired-TestDist}
+\raggedright
+\includegraphics[scale=0.14]{P-dist-subjects-big}
+
+\justifying
+\begin{multicols}{2}
+
+Each observation in the data we used was a specific experimental observation
+resulting of an individual transcranial stimulation. The original features of the data
+were:
+
+\begin{enumerate}
+    \item An $EMG$ variable with the EMG peak-to-peak of each observation.
+    \item A $Label$ categorical feature, which encoded the group of
+        the subject of each observation, was the target variable. 
+    \item An $ISI$ feature that econded the inter-stimulus interval of each
+        pulse. A value of $-1$ indicated that the given pulse was a test pulse (no
+        inter-stimulus interval).
+\end{enumerate}
+
+To these features, we included (separately and then together) the engineered
+features $\rho$ and $\delta$, which were computed for each observation using the
+Julia programming language. The code used to compute the features is publicly
+available (at \ldots).
+
+\section{Empirical results}
+
+The objective was to evaluate whether the inclusion of the engineered,
+pulse-specific measures of relative amplitude improved the performance of a
+machine learning model, and in what degree. In order to do this, we set a
+random forests model with the task of classifying every observation with its
+appropriate label. In order words, the model was set out to infer, based on the
+properties of each transcranial stimulation response, the group of the subject
+upon which the stimulation was elicited, as well as the type of sleep session
+after which it occurred.
+
+We trained the model first on the data without pulse-specific
+measures of relative amplitude (this is, with only the three features listed in
+the \textbf{Statistics} section). We then trained the exact same model on the
+data with the inclusion of $\delta$ but not $\rho$, and the inclusion of $\rho$
+but not $\delta$. We then trained the same model with both engineered features,
+and lastly with both features and also their weighted variantes, using
+inverse-variance weights.
 
@@
 
-When trained on the raw data, without the inclusion of
-pulse-specific relative amplitude measures, the model's
-accuracy was $A\approx34.2 \%$.
+When trained on the raw data, without the inclusion of pulse-specific relative
+amplitude measures, the model's accuracy was of $\approx34.2 \%$. However, the
+confusion of the model shows the errors are concentrated on the healthy control
+categories, while categorization of diagnosed subjects was substantially better.
+This implies major depressive subjects show statistically significant and
+distinct patterns in their TMS responses in comparison to healthy controls, and
+may be considered evidence for depression-induced differences in
+neuroplasticity.
 
-\pagebreak
+\end{multicols}
+
+\centering
+~
+
+\includegraphics[scale=0.4]{cm-raw}
+
+\justifying
+\begin{multicols}{2}
+
+However, it is important to note that the proxy for neuroplasticity in the
+double pulse paradigm is precisely the relative amplitude of double pulses with
+respect to test pulses. The raw data lacks any measure of relative amplitude.
+Besides, regardless of the fact that the model above suggests a critical
+difference in the response patterns between subject groups, its accuracy is
+still poor.
+
+\subsection{Engineered data}
+
+With the inclusion of the $\rho$ feature, accuracy increases by a factor of
+$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature alone increased
+it, in comparison to the raw data, had a more or less equivalent impact,
+increasing accuracy to $73.4\%$.
+
+\end{multicols}
+\raggedright
+
+~
+
+\includegraphics[scale=0.24]{cm-rho}
+\raggedright
+\includegraphics[scale=0.24]{cm-delta}
+
+\justifying
+\begin{multicols}{2}
+
+The model still shows a higher accuracy when classifying diagnosed subjects, but
+the overall accuracy across all categories was significantly improved.
+
+As stated earlier, the inclusion of the $\delta$ and $\rho$ features computed
+over the logarithmically transformed EMG peak-to-peak improved the model's
+accuracy. Concretely, accuracy was increased to $86\%$, with the following
+confusion matrix.
+
+\end{multicols}
+
+~
+\centering
+
+\includegraphics[scale=0.4]{cm-dr-ln}
+
+\justifying
+\begin{multicols}{2}
+
+At last, if to the previous model we add also the weighted versions of $\delta$
+and $\rho$, using inverse-variance weights, we obtained an accuracy of $89.8\%$,
+with the following confusion matrix.
+
+\end{multicols}
+
+\centering
+~
+
+\includegraphics[scale=0.4]{final-cm}
+
+\justifying
+\begin{multicols}{2}
+
+\section{Discussion}
+
+The previous results show that the inclusion of pulse-specific relative
+amplitude measures greatly improve the accuracy of a random forests model
+trained over TMS data for subject classification. More importantly, it becomes
+clear that, via the engineered features $\delta$ and $\rho$, researchers can
+attain evidence in favour or against hypotheses pertaining to differences among
+subject groups. Particularly, in showing that pulse-specific measures of
+relative amplitude are significant information to determine the sleep session
+and group of a subject, and being relative amplitude measures a proxy for
+neuroplasticity, the previous model provides evidence in favour of
+sleep-modulated, depression-induced differences in neuroplasticity.
+
+The use of machine learning models over neuroscientific observations is not only
+a promising tool in the production of evidence. There is a diagnostic potential
+that is still to be evaluated. Indeed, if machine learning models can detect the
+neural patterns that distinguish, under specific experimental conditions,
+healthy control from diagnosed subjects, such models can potentially be
+implemented in the diagnosis process as powerful clinical tool.
+
+Although long and serious scientific effort is still required to appraise the
+diagnostic potential of machine learning models, we believe our results allow
+for a certain amount of conservative optimism on the matter.
+
+In short, pulse-specific relative amplitude features are successful in making
+machine learning models applicable to TMS experimental data. Thus, they can play
+an important part in future research by allowing for new ways of analyzing and
+extracting meaningful information of TMS results.
+
 \section{Appendix}
 
-\subsection*{Proofs}
+\textbf{Proof 1}. In \textbf{Remark 1}, we observed the following property:
 
-\begin{theorem}
-    $\forall x: x \in \mathbb{R}^+ : \delta(x) \geq \rho(x)$
-\end{theorem}
+    \begin{align*}
+        \forall x: x \in \mathbb{R}^+ : \delta(x) \geq \rho(x)
+    .\end{align*}
 
-Recall that
+Such property can be proven via induction. Firstly, recall that
 
@@
 This holds for $m=1$, since $\frac{1}{t_1}+t_1 \geq 1 \iff 1+t_1^2 \geq t_1$. So
-we may assume $S^k_1 S^k_2 \geq k^2$ . We may only show now that
+we may assume $S^k_1 S^k_2 \geq k^2$ . We now set out to show that
 

### 2023-04-14 jl/analysis.jl

@@ -1,7 +1,9 @@
 using DataFrames
 using CSV
 using Statistics
-include("math.jl")
+using Plots
+using StatsPlots
+include("jl/math.jl")
 
 function sra(pulses, subject, session)
     """Computes standard relative amplitude given
@@ -32,3 +34,24 @@ function group_analysis(an, col=:SRA)
     an = combine(an, [col] => x -> mean(x))
     rename!(an, 3 => :SRA)
 end
+
+df = CSV.read("LnDF.csv", DataFrame)
+display(df)    
+
+s_df = subject_analysis(df)
+display(s_df)
+g_df = group_analysis(s_df)
+display(df)
+
+
+set_ρ(df)
+set_ρ_ln(df)
+set_ρ₂(df)
+set_ρ₂_ln(df)
+df = compute_inverse_variance_weights(df)
+set_weighted_ρ()
+set_weighted_ρ₂()
+
+display(df)
+
+CSV.write("full_df.csv", df)

### 2023-04-14 jl/research.jl

@@ -23,9 +23,7 @@ function compute_inverse_variance_weights(df)
     to each pulse in the data."""
     t = groupby(df, [:ISI, :Label])
     df = transform(t, :EMGPeakToPeak .=> inv_var_weights => :InvVarWeights)
-end
-
-# Helper functions
+end # Helper functions
 
 function get_pulses(subject, session, isi, colname=:EMGPeakToPeak)
     """Returns some column of the data filtered by subject, session, and isi. 
@@ -87,14 +85,32 @@ end
 function set_ρ(df)
     """Sets raw, pulse-specific relative amplitude feature in the database."""
     gdf = groupby(df, [:ISI, :Subject, :Session])
+    f(x, y, z) = ρ(x, get_test_pulses(df, y, z))
     res = transform(gdf, [:EMGPeakToPeak, :Subject, :Session] => (x, y, z) -> ρ(x, get_test_pulses(df, y, z)))
     df[!, :RRA] = res[:, end]
+end
+
+function set_ρ_ln(df)
+    """Sets raw, pulse-specific relative amplitude feature in the database."""
+    gdf = groupby(df, [:ISI, :Subject, :Session])
+    f(x, y, z) = ρ(log.(x), log.(get_test_pulses(df, y, z)))
+    res = transform(gdf, [:EMGPeakToPeak, :Subject, :Session] => (x, y, z) -> f(x, y, z))
+    df[!, :LogRRA] = res[:, end]
 end
 
 
 function set_ρ₂(df)
     """Sets raw, pulse-specific relative amplitude feature in the database."""
     gdf = groupby(df, [:ISI, :Subject, :Session])
-    res = transform(gdf, [:EMGPeakToPeak, :Subject, :Session] => (x, y, z) -> ρ₂(x, get_test_pulses(df, y, z)))
+    f(x, y, z) = ρ₂(x, get_test_pulses(df, y, z))
+    res = transform(gdf, [:EMGPeakToPeak, :Subject, :Session] => (x, y, z) -> f(x, y, z))
     df[!, :RRA2] = res[:, end]
+end
+
+function set_ρ₂_ln(df)
+    """Sets raw, pulse-specific relative amplitude feature in the database."""
+    gdf = groupby(df, [:ISI, :Subject, :Session])
+    f(x, y, z) = ρ₂(log.(x), log.(get_test_pulses(df, y, z)))
[…]

### 2023-04-14 jl/statistics.jl

@@ -1,0 +1,56 @@
+using DataFrames
+using CSV
+using Statistics
+using Plots
+using StatsPlots
+using Distributions
+
+include("jl/math.jl")
+
+# TEST AND NO-TEST DISTRIBUTIONS 
+
+qqplot(Exponential(248), df.EMGPeakToPeak)
+mean(df.EMGPeakToPeak)
+
+no_tests = filter(:ISI => x -> x != -1, df).EMGPeakToPeak
+tests = filter(:ISI => x -> x == -1, df).EMGPeakToPeak
+mean(no_tests)
+
+qqplot(Exponential(260), no_tests)
+
+density(no_tests, linewidth=2, color=:blue, label="Paired pulse distribution")
+density!(tests, linewidth=2, color=:red, label = "Test pulse distribution")
+xlabel!("EMG Peak to peak")
+ylabel!("Probability density")
+
+# GROUP - TYPE DISTRIBUTIONS 
+
+function get_group(group, type, tests)
+    data = filter([:Group, :Type] => (x, y) -> x == group && y == type, df)
+    if tests 
+        data = filter([:ISI] => x -> x == -1, data).EMGPeakToPeak
+    else
+        data = filter([:ISI] => x -> x != -1, data).EMGPeakToPeak
+    end
+    return data
+end 
+
+HCBL = get_group(1, "BL", false)
+HCSWD = get_group(1, "SWD", false)
+MDDBL = get_group(2, "BL", false)
+MDDSWD = get_group(2, "SWD", false)
+
+
+density(HCBL, linewidth=2, color=:blue, label="HC BL")
+density!(HCSWD, linewidth=2, color=:green, label = "HC SWD")
+density!(MDDBL, linewidth=2, color=:red, label="MDD BL")
+density!(MDDSWD, linewidth=2, color=:purple, label = "MDD SWD")
+title!("Paired pulse distribution across subject groups")
+xlabel!("EMG Peak to peak")
+ylabel!("Probability density")
+
+median(HCBL)
+median(HCSWD)
+median(MDDBL)
+median(MDDSWD)
+

### 2023-04-14 paper/paper.tex
Paper

@@
 \documentclass{article}
+\usepackage{subfigure}
 \usepackage{amsmath}
+\usepackage{float}
 \usepackage{ragged2e}
@@
 
-The emergence of new computational models has provided neuroscientific research
-with novel methods for extracting significant insights from experimental results.
-This paper takes a computational approach to challenges arising in the analysis
-of transcranial stimulation (TMS) data. We argue that conducting TMS experiments
-under the double-pulse paradigm enables the use of machine learning models. In
-particular, we formulate pulse-specific measures of relative amplitude between
-paired and test pulses. These has potential implications in both $a.$ the
-production of evidence and $b$. the diagnosis of clinical disorders.
+The emergence of new computational models has provided neuroscientists with
+novel ways of extracting significant insights from experimental data. This paper
+takes a computational approach to challenges arising from the analysis of
+transcranial magnetic stimulation (TMS) data under the double pulse paradigm. We
+argue that using indicators of neuroplasticity for each specific double
+stimulation, as opposed to their average, allows for the use of data-driven
+methods of analysis.
+
+Particularly, we formulate pulse-specific measures of relative amplitude between
+paired and test pulses. These measures represent the proportionality of specific
+paired-pulse responses with respect to test-pulse responses. Namely, they serve as
+proxies for neuroplasticity in the brain by capturing the degree of
+inhibition or facilitation induced by each double-stimulation.
+
+The benefits of data-driven methods of analysis cannot be overstated. Their
+potential implication in the production of scientific evidence and the
+diagnosis of clinical disorders go beyond the scope of this paper and deserve
+further investigation.
 
@@
 
-In the context of this paper, we shall term any coefficient
-that serves to represent the proportional relationship
-between the potentials evoked by paired and test pulses a
-\textit{measure of relative amplitude}.
+In the context of this paper, we shall term any coefficient that serves to
+represent the proportional relationship between the potentials evoked by paired
+and test pulses a \textit{measure of relative amplitude} (MRA).
 
@@
 individual paired pulse with respect to the set of test pulses in an
-experimental session. The purpose of this endeavor is to keep data resolution
-at its highest, which on its turn allows for the implementation of data-driven
-artificial intelligence in the analysis of the 
-experimental results. Thus, from a computer science perspective, our objective is
-constrained to the sphere of feature engineering.
+experimental session. The benefit of this is that it keeps data resolution at
+its highest, which allows for the implementation of data-driven artificial
+intelligence in the analysis of the experimental results. Thus, from a computer
+science perspective, our objective is constrained to the sphere of feature
+engineering.
 
@@
 
-\section{Data collection}
+\section{Dataset}
 
-We used data collected across $N = ?$ subjects at the \textit{Laboratory for the
-Study of Slow-wave sleep activity}, University of Pennsylvania, with $H = ?$
-healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
-Transcranial magnetic stimulation of the motor cortex was elicited to them after
-a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
-Motor evoked potentials were measured via an electrode (?) placed on the
-subjects' thumb (?). In the slow-wave disruption session, an auditory stimulus with
-sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet
-not strong enough to wake the subjects, was elicited. This experimental setting
-produces four distinct categories, two depending on the subject group and two on
-the type of sleep session underwent, as shown in the table below.
+Our case study uses data from $N = 43$ subjects at UPenn's \textit{Lab for the
+Study of Slow-wave sleep activity}, with $H = 17$ healthy controls (HC) and $D =
+26$ with major depressive disorder (MDD). Subjects received transcranial
+magnetic stimulation of the motor cortex after both a night of baseline sleep
+and a night of slow-wave disruption (SWD) sleep. The stimulations were performed
+under the double pulse paradigm. Per session, subjects received twenty-four
+($24$) test pulses in total and eight ($8$) paired pulses per ISI. The ISIs were
+$4, 5, 8, 10, 15$ and $20$. The evoked potentials were measured by placing an
+EMG electrode on the targeted muscle of the thumb. In the slow-wave disruption
+session, we sent a constant auditory stimulus with sufficient strength to
+interrupt the normal occurrence of slow-wave sleep, yet not strong enough to
+wake the subjects.
+
+Please see the table below to see the resulting subject categories:
+
+~
 
@@
 Major depressive disorder & MDD BL & MDD SWD \\
-\hli
+\hline
 \end{tabular}
@@
 
-Each observation in the data was a specific experimental sample resulting
-from an individual transcranial stimulation. The original features of the data
-were:
+~
+
+Each observation comes from an individual transcranial stimulation. The features
+of the resulting database are:
 
 \begin{enumerate}
-    \item An $EMG$ variable with the EMG peak-to-peak of each observation.
-    \item A $Label$ categorical feature, which encoded the group of
-        the subject of each observation, was the target variable. 
-    \item An $ISI$ feature that econded the inter-stimulus interval of each
+    \item An $EMG$ variable with the peak-to-peak EMG of each elicited pulse.
+    \item A $Label$ categorical feature, which represents the group of
+        the subject of each observation. 
+    \item An $ISI$ feature that represents the inter-stimulus interval of each
         pulse. A value of $-1$ indicated that the given pulse was a test pulse (no
@@
 
-\section{Limitations of the traditional analysis}
-
-Measures of relative amplitude in neuroscience are generally computed at the
-subject level by repeatedly taking averages in the following manner. First, the
-quotient of the average paired-pulse and the average test pulse is found per
-every subject. Secondly, those quotients are averaged out across all subjects in
-a subject group. This is reasonable, since hypotheses generally deal with
-differences across subject groups.
-
-Two major limitations arise from this method. The principle one is the fact that
-it implies greatly downscaling the data resolution Secondly, potentials evoked
-by test and paired pulses present a very high variance (Rossini et. al., 20015;
-Orth, Snijders and Rothwell, 2003; Wassermann, 2002), which —being that it is a
-quotient of two means—  raises questions about the quality of the representation
-where relatively few pulses are elicited.
-
-When considering the size of the data across multiple experimental subjects,
-this shrinkage in data resolution is far from being negligible. And in times
-when scientific inquiry can greatly benefit from forms of artificial
-intelligence that rely on large amounts of data (e.g. machine learning), it is
-clearly limiting to only count with low-resolution measures to interpret
-experimental results.
-
-\section{Proposed approach}
-
-\subsection{Pulse-specific relative amplitude measures}
-
-Considering the limitations of the traditional approach, this paper proposes two
-pulse-specific relative amplitude measures. This allows for otherwise unfeasible
-computational analyses of TMS data, such as the use of machine learning (ML) models
-for the detection of different pulse-response patterns among different groups of
-clinical subjects.
-
-To empirically test the validity of these measures, we evaluated whether they
-improved the performance of a classification model, and in what degree. (We
-chose a classification model for practical reasons --namely, because the
-samples in our data were conveniently labeled by subject group and sleep
-session).
-
-In order to do this, we defined a random forest model in Python under the
-following parameters: \ldots
-
-The model was tasked with classifying every observation with the appropriate subject
-label. In other words, it was set out to infer, based on the properties of each
-transcranial stimulation response, the group of the subject upon which the
-stimulation was elicited, as well as the type of sleep session after which it
-occurred.
-
-The model was first trained on the raw data, with the features described in
-\textbf{3.1}. Then, it was trained with the inclusion of one or many engineered
-features of relative amplitude (see section \textbf{4}). These features were
-computed for each observation using the Julia programming language.
-
-Since the neural response to transcranial stimulations follows an exponential
-distribution (see \textbf{Cite section}), we experimented with the inclusion of
-the features above applied to the logarithmically transformed peak-to-peak EMG.
-
-As stated earlier, action potentials evoked by transcranial magnetic
-stimulations follow a Gamma distribution that is very close to the exponential.
-(see \textbf{Statistics} section). Experimentation has shown that the inclusion
-of weighted variants of the features defined above greatly improves the
-performance of a random forests model (see \textbf{Empirical results}).
-
-The use of weighted instead of arithmetic averages may be useful in dealing with
-the excessive influence of outliers or highly spread out points in the feature.
-For example, the weight vectors may be computed using the MAD or the
-inverse-variance of each $x$.
-
-Thus, our hypothesis is that pulse-specific measures of relative
-amplitude will significantly improve a random forest's performance.
-
-The following section (\textbf{Cite section}) defines the proposed measures of
-relative amplitude. In section \textbf{Results}, we show the impact of these
-measures on the performance of the random forest model.
-
-\subsection{Definitions: Pulse specific relative amplitudes}
-
-For simplicity, we will deal with the hypothetical situation
-in which a single ISI was used for paired stimulations. All
-of our results generalize to different inter-stimulus intervals.
-
-Let $k$ be the number of experimental subjects in some
-subject group $\mathcal{G}$, to each of whom $n$ paired stimulations and
-$m$ test stimulations were elicited.
-
-\begin{definition}
-
-Let $\textbf{P}^{n \times k}, \textbf{T}^{m\times k}$ be matrices over the
-vector space $\mathbb{R}^+$ representing
-the paired and test potentials evoked across each of the $k$ subjects.
-
-\end{definition}
-
-\begin{definition} 
-    Let $x \in \textbf{P}_{\star i}$ be the MEP of a single paired stimulation
-    elicited on the $i$th experimental subject, and $\textbf{t} =
-    \textbf{T}_{\star i}$ the vector containing the MEPs of all test pulses
-    elicited on that subject. Let $\textbf{w}$ be some appropriate weight
-    vector. Then we define two pulse-specific relative amplitude measures,
-
-    \begin{equation} 
-        \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
-    \end{equation}
-
-    \begin{equation} 
-        \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} 
-    \end{equation}
-
-    \begin{equation}
-            \rho_w(x) &:= \frac{xm\sum_{j=1}^m w_j}{\sum_{j=1}^m t_j w_j} \\ 
-    \end{equation}
-
-    \begin{equation}
-        \delta_w(x) &:= \frac{\frac{x}{m}\sum_{j=1}^m\frac{w_j}{t_j}}{\sum_{j=1}^mw_j} 
-    \end{equation}
-\end{definition}
-
-\begin{remark} 
-    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
-    property, consult the appendix.) 
-\end{remark}
-
-Notice that $\rho(x)$ is the proportion between the potential $x$, evoked by a
-paired stimulation, with respect to the average potential of single test
-stimulations. On the other hand, $\delta(x)$ is the average proportion of $x$
-with respect to each single test pulse.
-
-Notice as well that $\rho$ and $\delta$ are defined over the test pulses in a
-specific experimental session $i$. In other words, whenever we speak of the
-$\delta$ and $\rho$ \textit{features} (\textit{id est}, of the data resulting from
-broadcasting $\rho$ and $\delta$ over a series of observations), such feature is
-understood to be subject- and session-specific.
-
-In particular, $\rho$ is a representation of the relative importance of each
-double pulse in relation to the overall distribution of the test pulses in the
-subject. It measures the deviation of the double pulse $x$ with respect to the
-average test pulse. On the other hand, $\delta$ is a measure of the
-proportionality of a double pulse with respect to different values in the
-distribution of the test pulses of the subject.
-
-\section{Results}
-
-\subsection{ML model without relative amplitude measures}
-
-When trained on the raw data, without the inclusion of pulse-specific relative
-amplitude measures, the model's accuracy was of $\approx34.2 \%$. However, the
-confusion matrix of the model shows the errors are concentrated on the healthy
-control categories, while categorization of diagnosed subjects was substantially
-better. This implies major depressive subjects show statistically significant
-and distinct patterns in their TMS responses in comparison to healthy controls,
-and may be considered evidence for depression-induced differences in
-neuroplasticity.
-
-\end{multicols}
-
-\centering
-~
-
-\includegraphics[scale=0.4]{cm-raw}
-
-\justifying
-\begin{multicols}{2}
-
-However, it is important to note that the proxy for neuroplasticity in the
-double pulse paradigm is precisely the relative amplitude of double pulses with
-respect to test pulses. The raw data lacks any measure of relative amplitude.
-Besides, regardless of the fact that the model above suggests a critical
-difference in the response patterns between subject groups, its accuracy is
-still poor.
-
-\subsection{Engineered data}
-
-With the inclusion of the $\rho$ feature, accuracy increases by a factor of
-$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature alone increased
-it, in comparison to the raw data, had a more or less equivalent impact,
-increasing accuracy to $73.4\%$.
-
-\end{multicols}
-\raggedright
-
-~
-
-\includegraphics[scale=0.24]{cm-rho}
-\raggedright
-\includegraphics[scale=0.24]{cm-delta}
-
-\justifying
-\begin{multicols}{2}
-
-The model still shows a higher accuracy when classifying diagnosed subjects, but
-the overall accuracy across all categories was significantly improved.
-
-As stated earlier, the inclusion of the $\delta$ and $\rho$ features computed
-over the logarithmically transformed EMG peak-to-peak improved the model's
-accuracy. Concretely, accuracy was increased to $86\%$, with the following
-confusion matrix.
-
-\end{multicols}
-
-~
-\centering
-
-\includegraphics[scale=0.4]{cm-dr-ln}
-
-\justifying
-\begin{multicols}{2}
-
-At last, if to the previous model we add also the weighted versions of $\delta$
-and $\rho$, using inverse-variance weights, we obtained an accuracy of $89.8\%$,
-with the following confusion matrix.
-
-\end{multicols}
-
-\centering
-~
-
-\includegraphics[scale=0.4]{final-cm}
-
-\justifying
-\begin{multicols}{2}
-
-\section{Statistics}
-
-\begin{wrnote}
-    The description of the experimental process in this section is poor. Also I
-    do not know the total number of subjects, experiments are still being
-    conducted. Need input from the lab on these matters.
-\end{wrnote}
-
-Although distributions were always exponential, the $\beta$ parameter of said
-distributions varied across subject groups and session types.
+The distribution of the evoked potentials was always exponential. However, the
+$\beta$ parameter of the distribution varies across subject groups and session
+types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2}.
 
@@
 
-Each of the means in the above table correspond to the estimated $\beta$
+Each of the means in the above table corresponds to the estimated $\beta$
 parameter of the distribution of the paired pulses on each subject group.
 
-\end{multicols}
-\raggedright
+\begin{figure*}
+  \begin{minipage}[b]{0.48\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{Paired-TestDist}
+    \caption{Distribution of paired and test pulses}
+    \label{fig:figure1}
+  \end{minipage}
+  \hspace{0.04\linewidth}
+  \begin{minipage}[b]{0.48\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{P-dist-subjects-big}
+    \caption{Distribution of paired pulses across subject groups}
+    \label{fig:figure2}
+  \end{minipage}
+\end{figure*}
 
-~
+\section{Limitations of the traditional TMS method of analysis}
 
-\includegraphics[scale=0.14]{Paired-TestDist}
-\raggedright
-\includegraphics[scale=0.14]{P-dist-subjects-big}
+Measures of relative amplitude in neuroscience are generally computed at the
+subject and group level by repeatedly taking averages. This is natural, since
+scientific interest generally resides in evaluating differences across subject
+groups. However, it also raises problems pertaining to
+the downscaling of the data resolution and the reliability of the resulting values.
 
-\justifying
-\begin{multicols}{2}
+The procedure to establish the average relative amplitude of a group of subjects
+(for example, healthy controls) is simple. Traditionally, it involves taking the
+quotient of the average paired-pulse and the average test pulse responses per
+every subject. Next, those quotients are averaged out across all subjects in the
+subject group. In short, it is an average of averages.
+
+Although this practice is not arbitrary ---since hypotheses generally deal with differences
+across subject groups--- it has three major limitations. Firstly, it implies
+greatly downscaling the data resolution, making data-driven methods of analysis
+unfeasible. Secondly, potentials evoked by test and paired pulses present a very
+high variance (Rossini et. al., 20015; Orth, Snijders and Rothwell, 2003;
+Wassermann, 2002), which —being that it is a quotient of two means—  raises
+questions about the quality of the representation where relatively few pulses
+are elicited. Thirdly, our observations suggest potentials evoked by TMS
+stimulations follow an exponential distribution, which puts in question whether
+arithmetic means are a reliable measure of central tendency.
+
+To make matters worse, when considering the size of the data across multiple
+experimental subjects, the shrinkage in data resolution is far from negligible.
+And in times when scientific inquiry can greatly benefit from forms of
+artificial intelligence that rely on large amounts of data (e.g. machine
+learning), it is clearly limiting to count solely with low-resolution measures
+to interpret experimental results.
+
+\section{Proposed approach}
+
+\subsection{Pulse-specific relative amplitude measures}
+
+In light of the technical concerns about the traditional approach, this paper
+proposes two pulse-specific relative amplitude measures. This allows for
+otherwise unfeasible computational analyses of TMS data, such as the use of
+machine learning (ML) models for the detection of different pulse-response
+patterns among different groups of clinical subjects.
+
+Pulse-specific measures of relative amplitude are coefficients that represent,
+for every experimental subject, the proportionality of each individual
+paired-pulse response with respect to the set of \textit{all} test-pulse
+responses. From a neuroscientific perspective, they are a proxy for
+neuroplasticity at the level of each particular stimulation, measuring the
+amount facilitation or inhibition in each specific response. From a
+computational perspective, these measures prescind from the need of averaging
+paired-pulses out, and thus keep data resolution at its highest.
+
+In order to test these measures, we look at whether they improve the performance
+of a ML classification model and in what degree.\footnote{We use a
+classification model for practical reasons ---namely, because the samples in our
+data were conveniently labeled by subject group and sleep session.} Our case
+study is a random forest model in Python tasked with classifying every
+observation with the appropriate subject label. In other words, its goal is to
+infer, based on the properties of each TMS response, the group of the subject
+upon which the stimulation was elicited, as well as the type of sleep session
+after which it occurred. Our hypothesis is that pulse-specific measures of relative
+amplitude will significantly improve a random forest's performance.
+
+Section \textbf{5} provides the mathematical definitions of our measures of
+relative amplitude. Section \textbf{6} discusses the impact of each measure on
+the performance of the random forest model. Specifically, section \textbf{6.1}
+shows the performance of the model when trained over the raw data (i.e. the
+features described in section \textbf{3.1}). Section \textbf{6.2} shows the
+performance of the model when including pulse-specific measures of relative
+amplitude. Section \textbf{6.3} shows the performance of the model when combining the
+pulse-specific measures with logarithmic transformations of the raw data and the
+use of weighted instead of arithmetic averages.
+
+As for the computation of these measures, we implemented the formulas of section
+\textbf{5.2} for every TMS response using the Julia programming language.
+
+\subsection{Definitions: Pulse specific relative amplitudes}
+
+For simplicity, we deal with the hypothetical situation in which a single ISI is
+used for paired stimulations in a unique TMS session. Our formulations
+generalize easily to different inter-stimulus intervals and multiple
+experimental sessions.
+
+Firstly, we need to establish some convenient notation. Let $n$ and $m$ be the
+number of paired and test pulses elicited to the subject in the session,
+respectively. Let $\textbf{x}$, $\textbf{t}$ be vectors over the $\mathbb{R}^+$
+vector space whose components are the evoked potentials of paired and pulse
+stimulations, respectively.
+
+\begin{definition}
+
+    Let $x \in \textbf{x}$ be the evoked potential of a single paired
+    stimulation. Then we define two pulse-specific relative amplitude measures,
+
+    \begin{equation} 
+        \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
+    \end{equation}
+
+    \begin{equation} 
+        \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} 
+    \end{equation}
+\end{definition}
+
+\begin{remark} 
+    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
+    property, consult the appendix.) 
+\end{remark}
+
+Notice that $\rho(x)$ is the proportion between the potential $x$, evoked by a
+paired stimulation, with respect to the average potential of single test
+stimulations. On the other hand, $\delta(x)$ is the average proportion of $x$
+with respect to each single test pulse.
+
+To put it differently, $\rho$ is a representation of the relative importance of
+each double pulse in relation to the overall distribution of the test pulses in
+the subject. It measures the deviation of the double pulse $x$ with respect to
+the average test pulse. $\delta$, in contrast, measures the proportionality of a
+double pulse with respect to different values in the distribution of the test
+pulses of the subject.
+
+Notice as well that $\rho$ and $\delta$ are defined over the evoked potentials
+of a specific experimental session. In other words, whenever we speak of the
+$\delta$ and $\rho$ \textit{features} (\textit{id est}, of the data resulting
+from broadcasting $\rho$ and $\delta$ over a collection of evoked potentials),
+such features are understood to be subject- and session-specific.
+
+\section{Results}
+
+\subsection{Random forest without MRA}
+
+When trained on the raw data ---that is, without the inclusion of pulse-specific
+MRA--- the model's accuracy was of $\approx34.2 \%$. However, the confusion
+matrix of the model, \textbf{Figure 3}, shows the errors are concentrated on the healthy control
+categories, while categorization of diagnosed subjects was substantially better.
+This implies major depressive subjects show statistically significant and
+distinct patterns in their TMS responses in comparison to healthy controls, and
+may be considered evidence for depression-induced differences in
+neuroplasticity.
+
+However, it is important to note that the proxy for neuroplasticity in the
+double pulse paradigm is precisely the relative amplitude of double pulses with
+respect to test pulses. The raw data lacks any measure of relative amplitude.
+Besides, regardless of the fact that the model above suggests a critical
+difference in the response patterns between subject groups, its accuracy is
+still poor.
+
+\subsection{Random forest with MRA}
+
+We then trained the model first with the aggregation of the $\rho$ feature, then
+with the aggregation of the $\delta$ feature, and then with the aggregation of
+both.
+
+With the inclusion of the $\rho$ feature increased accuracy by a factor of
+$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature had a similar
+impact, increasing accuracy to $73.4\%$.
+
+\begin{figure}[H]
+  \centering
+  \subfigure[Raw data]{\includegraphics[width=0.48\textwidth]{cm-raw}}
+  \hfill
+  \subfigure[$\rho$ and $\delta$ features included.]{\includegraphics[width=0.48\textwidth]{cm-rho-delta}}
+  \caption{Compared performance: raw data and data with measures of relative
+  amplitude.}
+  \label{fig:two_images}
+\end{figure}
+
+The inclusion of both the $\delta$ and $\rho$ features computed over the
+logarithmically transformed EMG peak-to-peak improved the model's accuracy.
+Concretely, accuracy was increased to $86\%$, with the following confusion
+matrix.
+
+\subsection{Random forest after $\ln$ transforms and weighted averages}
+
+Since the neural response to transcranial stimulations follows an exponential
+distribution (see \textbf{Cite section}), we experimented with the inclusion of
+the features above applied to the logarithmically transformed peak-to-peak EMG.
+
+As stated earlier, action potentials evoked by transcranial magnetic
+stimulations follow a Gamma distribution that is very close to the exponential.
+(see \textbf{Statistics} section). Experimentation has shown that the inclusion
+of weighted variants of the features defined above greatly improves the
+performance of a random forest model (see \textbf{Empirical results}).
+
+The use of weighted instead of arithmetic averages may be useful in dealing with
+the excessive influence of outliers or highly spread out points in the feature.
+For example, the weight vectors may be computed using the MAD or the
+inverse-variance of each $x$.
+
+As stated in section \textbf{3}, we experimented with using weighted
+versions of our pulse-specific MRA. They consisted in weighting test pulses
+with some appropriate weight vector. For example, including $\rho$ and
+$\delta$ using inverse-variance weights, we obtained an accuracy of
+$90.2\%$, with the following confusion matrix.
+
+\begin{figure*}
+    \centering
+    \includegraphics[scale=0.4]{final-cm}
+    \caption{Confusion matrix: $\rho(x)$, $\delta(x), \rho(\ln x), \delta(\ln
+    x)$, $\rho_w(x)$ and $\delta_w$(x).}
+    \label{fig:figure5}
+\end{figure*}
 
@@
 
-The previous results show that the inclusion of pulse-specific relative
-amplitude measures greatly improve the accuracy of a random forests model
-trained over TMS data for subject classification. More importantly, it becomes
-clear that, via the engineered features $\delta$ and $\rho$, researchers can
-attain evidence in favour or against hypotheses pertaining to differences among
-subject groups. Particularly, in showing that pulse-specific measures of
-relative amplitude are significant information to determine the sleep session
-and group of a subject, and being relative amplitude measures a proxy for
-neuroplasticity, the previous model provides evidence in favour of
-sleep-modulated, depression-induced differences in neuroplasticity.
+The results in \textbf{Section 6} demonstrate that pulse-specific MRA greatly
+improve the accuracy of a random forest model trained over TMS data for subject
+classification. More importantly, it becomes clear that, by using the engineered
+features $\delta$ and $\rho$, researchers can attain evidence in favour or
+against hypotheses pertaining to differences among subject groups. Particularly,
+the model provides otherwise unattainable evidence in favour of sleep-modulated,
+depression-induced differences in neuroplasticity.
 
-The use of machine learning models over neuroscientific observations is not only
-a promising tool in the production of evidence. There is a diagnostic potential
-that is still to be evaluated. Indeed, if machine learning models can detect the
-neural patterns that distinguish, under specific experimental conditions,
-healthy control from diagnosed subjects, such models can potentially be
-implemented in the diagnosis process as powerful clinical tool.
+\section{Conclusion}
+
+This paper shows that pulse-specific relative amplitude measures allow for the
+use of making machine learning models over TMS experimental data. Such models
+can play an important part in future research by introducing new ways of
+both analyzing and extracting meaningful insights from TMS results.
+
+The use of machine learning models in neuroscientific observations is not only a
+promising tool in the production of evidence. There is a diagnostic potential
+that is yet to be evaluated. By detecting distinct neural patterns between
+healthy control and diagnosed subjects,\footnote{At least under specific
+experimental conditions.} these models can potentially serve as a useful tool in
+the diagnostic process.
 
 Although long and serious scientific effort is still required to appraise the
-diagnostic potential of machine learning models, we believe our results allow
-for a certain amount of conservative optimism on the matter.
+diagnostic potential of these technologies, we believe our results allow
+for cautious optimism on the matter.
 
-In short, pulse-specific relative amplitude features are successful in making
-machine learning models applicable to TMS experimental data. Thus, they can play
-an important part in future research by allowing for new ways of analyzing and
-extracting meaningful information of TMS results.
+It should however be noted that, although machine learning models reveal
+distinct response patterns and differences across subjects, they do not disclose
+the neurobiology modulating those differences nor what those patterns are.
 
-\section{Appendix 1}
+Lastly, the impact of pulse-specific MRA in data-driven models other than random forests
+is yet to be tested. These empirical trials are not trivial because failure or
+success of the models is dependent not only on the engineered features, but on
+the nature of the experimental data itself.
 
-\section{Appendix 2}
+\section{Appendix}
 

### 2023-04-14 paper/paper.tex
Paper

@@
 \newtheorem{proof}{Proof}
+
+\title{\textbf{Modelling Neuroplasticity: Using Machine Learning to Learn About TMS Data}}
+\author{Santiago López Pereyra, Jennifer Goldschmied, Coauthor 3,
+Coauthor 4, \ldots}
+\maketitle
 
@@
 $\beta$ parameter of the distribution varies across subject groups and session
-types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2}.
+types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2} and in the
+table below:
 
@@
     HC SWD & 165.98 & 99.76 \\
-    \hline \\ 
+    \hline 
     MDD BL & 187.56 & 127.69\\ 
@@
 
-Each of the means in the above table corresponds to the estimated $\beta$
-parameter of the distribution of the paired pulses on each subject group.
+Observe that each of the means in the above table corresponds to the estimated
+$\beta$ parameter of the distribution of the paired pulses on each subject
+group.
 
@@
 study is a random forest model in Python tasked with classifying every
-observation with the appropriate subject label. In other words, its goal is to
-infer, based on the properties of each TMS response, the group of the subject
-upon which the stimulation was elicited, as well as the type of sleep session
-after which it occurred. Our hypothesis is that pulse-specific measures of relative
-amplitude will significantly improve a random forest's performance.
+observation with the appropriate subject label. The model we use is an AdaBoost
+classifier consisting of 20 decision trees. Every tree in the ensemble has a
+maximum depth of $8$. The boosting algorithm is \textit{SAMME.R}. The learning
+rate is set to $\eta = 0.01$. We use holdout cross-validation to evaluate the
+performance of the model. The total number of values in our dataset is $n =
+9382$.
+
+The goal of the classifier is to infer, based on the properties of each TMS
+response, the group of the subject upon which the stimulation was elicited, as
+well as the type of sleep session after which it occurred. Our hypothesis is
+that pulse-specific measures of relative amplitude will significantly improve
+the model's performance.
 
@@
 
-When trained on the raw data ---that is, without the inclusion of pulse-specific
-MRA--- the model's accuracy was of $\approx34.2 \%$. However, the confusion
-matrix of the model, \textbf{Figure 3}, shows the errors are concentrated on the healthy control
-categories, while categorization of diagnosed subjects was substantially better.
-This implies major depressive subjects show statistically significant and
-distinct patterns in their TMS responses in comparison to healthy controls, and
-may be considered evidence for depression-induced differences in
-neuroplasticity.
+The first random forest model is our control group. We trained it on the raw
+data ---that is, without the inclusion of pulse-specific MRA. The model's
+accuracy on the testing set was of $\approx34.2 \%$. The confusion matrix of the
+model, \textbf{Figure 3.}\textit{a}, shows that errors concentrated towards
+healthy control categories. Additionally, the categorization of diagnosed
+subjects was poor, although substantially better than that of healthy controls.
 
-However, it is important to note that the proxy for neuroplasticity in the
-double pulse paradigm is precisely the relative amplitude of double pulses with
-respect to test pulses. The raw data lacks any measure of relative amplitude.
-Besides, regardless of the fact that the model above suggests a critical
-difference in the response patterns between subject groups, its accuracy is
-still poor.
+Although the model almost never confuses an MDD subject with a healthy control,
+this does not necessarily imply that the model is capturing key differences
+between these groups. On the contrary, this is more likely a consequence of the
+model's bias in favour of the MDD categories. This bias is probably caused by
+the fact that the dataset has more MDD subjects than healthy controls. In
+consequence, the model learns the rudimentary strategy of always guessing for
+one of the MDD categories.
 
@@
 
-We then trained the model first with the aggregation of the $\rho$ feature, then
-with the aggregation of the $\delta$ feature, and then with the aggregation of
-both.
-
-With the inclusion of the $\rho$ feature increased accuracy by a factor of
-$\approx 2.1$ to $72.6\%$. The inclusion of the $\delta$ feature had a similar
-impact, increasing accuracy to $73.4\%$.
+The following models incorporated the $\rho$ feature, then the $\delta$ feature,
+and then both, respectively. The inclusion of the $\rho$ feature increased
+accuracy by a factor of $\approx 2.1$ to $72.6\%$. Subsequently, the inclusion
+of the $\delta$ feature had a similar impact, increasing accuracy to $73.4\%$.
+Lastly, the inclusion of both the $\delta$ and $\rho$ features improved accuracy
+to $\approx 87.1\%$. The confusion matrix of the model trained with both $\rho$
+and $\delta$ can be observed in \textbf{Figure 3.}\textit{b}.
 
@@
 
-The inclusion of both the $\delta$ and $\rho$ features computed over the
-logarithmically transformed EMG peak-to-peak improved the model's accuracy.
-Concretely, accuracy was increased to $86\%$, with the following confusion
-matrix.
-
-\subsection{Random forest after $\ln$ transforms and weighted averages}
+\subsection{Random forest with additional features}
 
 Since the neural response to transcranial stimulations follows an exponential
-distribution (see \textbf{Cite section}), we experimented with the inclusion of
-the features above applied to the logarithmically transformed peak-to-peak EMG.
+distribution (see Figures 1 and 2), we tested the $\rho$ and $\delta$ functions
+over the logarithmically transformed peak-to-peak EMG. This had a negligible
+positive impact in the overall performance.
 
-As stated earlier, action potentials evoked by transcranial magnetic
-stimulations follow a Gamma distribution that is very close to the exponential.
-(see \textbf{Statistics} section). Experimentation has shown that the inclusion
-of weighted variants of the features defined above greatly improves the
-performance of a random forest model (see \textbf{Empirical results}).
+On the contrary, the inclusion of weighted variants of the features defined
+above greatly improves the performance of the model. The use of weighted instead
+of arithmetic averages in the formulas for $\delta$ and $\rho$ is useful in
+dealing with the excessive influence of outliers or highly spread out points in
+the feature. For example, the weight vectors may be computed using the MAD or
+the inverse-variance of each $x$.
 
-The use of weighted instead of arithmetic averages may be useful in dealing with
-the excessive influence of outliers or highly spread out points in the feature.
-For example, the weight vectors may be computed using the MAD or the
-inverse-variance of each $x$.
+For example, using inverse-variance weights accuracy increased to $\approx
+90.2\%$. The confusion matrix of this last model can be appreciated in
+\textbf{Figure 4}.
 
-As stated in section \textbf{3}, we experimented with using weighted
-versions of our pulse-specific MRA. They consisted in weighting test pulses
-with some appropriate weight vector. For example, including $\rho$ and
-$\delta$ using inverse-variance weights, we obtained an accuracy of
-$90.2\%$, with the following confusion matrix.
+\begin{figure}[H]
+    \centering
+    \subfigure[$\rho, \delta, \rho(\ln x), \delta(\ln x), \rho_w$ and $\delta_w$
+    included.]{\includegraphics[width=0.48\textwidth]{final-cm}}
+    \caption{Confusion matrix}
+    \label{fig:figure4}
+\end{figure}
 
-\begin{figure*}
-    \centering
-    \includegraphics[scale=0.4]{final-cm}
-    \caption{Confusion matrix: $\rho(x)$, $\delta(x), \rho(\ln x), \delta(\ln
-    x)$, $\rho_w(x)$ and $\delta_w$(x).}
-    \label{fig:figure5}
-\end{figure*}
-
-\section{Discussion}
-
-The results in \textbf{Section 6} demonstrate that pulse-specific MRA greatly
-improve the accuracy of a random forest model trained over TMS data for subject
+In summary, our results demonstrate that pulse-specific MRA greatly improve the
+accuracy of a random forest model trained over TMS data for subject
 classification. More importantly, it becomes clear that, by using the engineered
@@
 This paper shows that pulse-specific relative amplitude measures allow for the
-use of making machine learning models over TMS experimental data. Such models
-can play an important part in future research by introducing new ways of
-both analyzing and extracting meaningful insights from TMS results.
+use of machine learning models over TMS experimental data. Such models
+can play an important part in TMS research by introducing new ways of
+both analyzing and extracting meaningful insights from the experimental results.
 
@@
 the nature of the experimental data itself.
+
+\end{multicols}
+
+\section{References}
+
+\begin{enumerate}
+    \item Lefaucheur, JP. “Transcranial Magnetic Stimulation.” Handbook of clinical neurology. U.S. National Library of Medicine. Accessed April 14, 2023. https://pubmed.ncbi.nlm.nih.gov/31277876. 
+    \item Orth, M, AH Snijders, and JC Rothwell.
+“The Variability of Intracortical Inhibition and Facilitation.” Clinical neurophysiology : official journal of the International Federation of Clinical Neurophysiology.
+U.S.
+National Library of Medicine.
+Accessed April 14, 2023.
+https://pubmed.ncbi.nlm.nih.gov/14652096/.
+    \item Rossini, PM, D Burke, LG Cohen, Z Daskalakis, R Di Iorio, V Di Lazzaro, F Ferreri, et al.
+“Non-Invasive Electrical and Magnetic Stimulation of the Brain, Spinal Cord, Roots and Peripheral Nerves: Basic Principles and Procedures for Routine Clinical and Research Application.
+an Updated Report from an I.F.C.N.
+Committee.” Clinical neurophysiology : official journal of the International Federation of Clinical Neurophysiology.
+U.S.
+National Library of Medicine.
+Accessed April 14, 2023.
+https://pubmed.ncbi.nlm.nih.gov/25797650/.
+        \item Wassermann, EM. “Variation in the Response to Transcranial
+            Magnetic Brain Stimulation in the General Population.” Clinical
+            neurophysiology : official journal of the International Federation
+            of Clinical Neurophysiology. U.S. National Library of Medicine.
+            Accessed April 14, 2023. \newline https://pubmed.ncbi.nlm.nih.gov/12088713/. 
+\end{enumerate}
+
+\pagebreak
+
+\begin{multicols}{2}
 

### 2023-06-11 paper/paper.tex
.

@@
 
-The distribution of the evoked potentials was always exponential. However, the
+The distribution of the evoked potentials is always exponential. However, the
 $\beta$ parameter of the distribution varies across subject groups and session
@@
 
-For simplicity, we deal with the hypothetical situation in which a single ISI is
-used for paired stimulations in a unique TMS session. Our formulations
-generalize easily to different inter-stimulus intervals and multiple
-experimental sessions.
+For simplicity, we will first deal with the hypothetical situation in which a
+single ISI is used for paired stimulations in a unique TMS session. Our
+formulations generalize easily to different inter-stimulus intervals and
+multiple experimental sessions.
 
@@
 respectively. Let $\textbf{x}$, $\textbf{t}$ be vectors over the $\mathbb{R}^+$
-vector space whose components are the evoked potentials of paired and pulse
-stimulations, respectively.
+vector space whose components are understood to be the evoked potentials of
+paired and test stimulations, respectively.
 
@@
     stimulation. Then we define two pulse-specific relative amplitude measures,
+    $\rho : \mathbb{R} \to \mathbb{R}, \delta : \mathbb{R} \to \mathbb{R}$,
+    where
 
@@
 such features are understood to be subject- and session-specific.
+
+\begin{remark}
+    Let $\pi$ denote the coefficient traditionally used for estimating the
+    relative amplitude between test and paired pulses. In accordance to what was
+    described in the second paragraph of Section $4$,
+
+    \begin{align*}
+        \pi = \frac{\mu_x}{\mu_t}
+    .\end{align*}
+
+    where $\mu_x, \mu_t$ are the mean paired and test pulse responses,
+    respectively. It follows from the definition of $\pi$ and $\rho$ that
+
+    \begin{align*}
+        \pi = \frac{\rho(x_1) + \ldots + \rho(x_n)}{n}
+    .\end{align*}
+
+    (For a proof, consult the appendix.)
+\end{remark}
+
+The second remark is of great importance. It states that the way in which
+relative amplitude was traditionally computed---the way in which neuroplasticity
+was traditionally proxied---is different from our proposed $\rho$ function only
+in terms of scale. More specifically, the traditional coefficient used to proxy
+neuroplasticity at the subject level is nothing but the mean $\rho$
+measure in an experimental session. We believe this bestows $\rho$ with a
+certain degree of confidence, insofar as not only it is not entirely new, but
+neuroscientists have implicitly used it all along.
+
+\subsection{Generalizing to many subjects}
+
+\begin{definition}
+    Let $\textbf{x} \in \mathbb{R}^j$. Then we define $T(\textbf{x}) := \begin{bmatrix} 
+        \rho(x_1) & \ldots & \rho(x_j)
+    \end{bmatrix}^\intercal $ and $R(\textbf{x}) := \begin{bmatrix} 
+        \delta(x_1) & \ldots & \delta(x_j) 
+    \end{bmatrix}^\intercal $.
+\end{definition}
+
+\begin{theorem}
+    The mappings $T, R$ are linear transformations.
+\end{theorem}
+
+Let $k \in \mathbb{N}$ denote the number of experimental subjects in an
+experiment, to each of whom $n, m$ paired and test pulses were elicited,
+respectively.
+
+\begin{definition}
+    We define $\textbf{P} \in \mathbb{R}^{k \times n}$ a
+    matrix where the $i$th row vector $\textbf{x}^{(i)}$ is the vector whose
+    components are the paired-pulse responses of the $i$th subject.
+\end{definition}
+
+\begin{definition}
+    Let $\mu^{(t)}_i$ be average test-pulse response of the $i$th subject. Then
+    we define
+    $\textbf{M}^{k \times k}$ a diagonal matrix with $\textbf{M}_{(ii)} =
+    \frac{1}{\mu^{(t)}_{i}}$.
+\end{definition}
+
+\begin{theorem}
+    The row-vectors of the matrix $\textbf{X}_\rho = \textbf{M}\textbf{P}$ are the
+    vectors $T(\textbf{x}^{(1)}), \ldots, T(\textbf{x}^{(k)})$.
+\end{theorem}
+
+Similarly, if we let $\textbf{M'}$ be a diagonal matrix such that
+$\textbf{M'}_{ii} = \mu^{(t^{-1})}_i$, where $\mu^{(t^{-1})}_i$ denotes the
+average \textit{inverse} test-pulse response of the $i$th subject, we come to
+the following result.
+
+\begin{theorem}
+    The row-vectors of the matrix $\textbf{X}_\delta = \textbf{M'}\textbf{P}$ 
+    are the vectors $R(\textbf{x}^{(1)}), \ldots,
+    R(\textbf{x}^{(k)})$.
+\end{theorem}
+
+If experiments were conducted using multiple ISI, then to each ISI $r$ it
+corresponds a distinct matrix $\textbf{X}_\delta^{ISI = r},
+\textbf{X}_\rho^{ISI = r}$. The same applies to different subject groups.
+
+We have expanded $\rho, \delta$ to higher dimensions for the purpose of
+providing a theoretical framework of their application to larger data sets and
+many experimental subjects. However, it is important to note that the matrix
+representations here presented only make sense under the assumption that, on
+each subject, none of $n$ paired pulses were anomalous, and hence all can be
+safely included in the model. Such assumption rarely holds in reality, where it
+is usual to excluded some pulses due to the presence of artifacts. There are
+many ways to deal with this situation in the implementation of our model---the
+safest being the direct application of $T$ and $R$ to the data vectors---. Thus,
+more than serving the algorithmic implementation of the model, which may
+prescind from them, these matrices provide insight into the nature of the
+transformations involved in the model.
 
@@
 the feature. For example, the weight vectors may be computed using the MAD or
-the inverse-variance of each $x$.
+the inverse-variance of each $t \in \textbf{t}$.
 
@@
 
+\textbf{Proof 2.} In \textbf{Remark 2}, we observe that $\rho$ is implicitly
+used in the traditional measures of relative amplitude. Recall that such methods
+consist in taking the average paired-pulse response and dividing it by the
+average test-pulse response. It is easily observed that
+
+\begin{align*} \frac{\mu_x}{\mu_d} = \frac{\frac{1}{n}\sum_{j=1}^n
+        x_{j}}{\frac{1}{m}\sum_{j = 1}^m t_j} 
+        = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
+\end{align*}
+
+\begin{proof}[Theorem 1]
+
+    Let $\sigma = \sum_{i=1}^m t_j$ and $c \in \mathbb{R}$. It follows directly
+    from the definition of $\rho$ that
+
+    \begin{align*}
+        T(c\textbf{x} + \textbf{y}) &= \begin{bmatrix} \frac{m(cx_1) + y_1}{\sigma} &
+        \ldots& \frac{m(cx_n + y_n)}{\sigma} \end{bmatrix}^\intercal \\ 
+                                    &=\begin{bmatrix} 
+                                        \frac{mcx_1}{\sigma} +
+                                        \frac{my_1}{\sigma} & \ldots &
+                                        \frac{mcx_n}{\sigma} + \frac{y_n}{\sigma} 
+                                    \end{bmatrix} \\
+                                    &=cT(\textbf{x}) + T(\textbf{y})
+    \end{align*}
+
+    The proof the $R$ is linear is practically identical.
+\end{proof}
+
 \end{document}

### 2023-06-11 R/loader.R

@@ -69,7 +69,7 @@ get_session_as_number <- function(session_name) {
 }
 
 # Horrible but serves its purpose and will
-# only be called once in human history.
+# only be called once in human history *cringes*
 get_label <- function(df) {
     if ((df$Group[1] == 1) && (df$Type == "SWD")) {
         return("HC SWD")
@@ -155,7 +155,6 @@ get_subject_data <- function(df, subject) {
     return(df)
 }
 
-
 create_isi_col <- function(df) {
     len <- nrow(df)
     vec <- c()
@@ -185,18 +184,18 @@ integrity_test <- function(df) {
     return(wrongs)
 }
 
-df <- load_data() %>% clean_data()
-write_csv(df, "df.csv")
-View(df)
-t <- read_delim("SWIP_018_S1.txt",
-    delim = "\t", comment = "#", na = c("", "NA", "(null)"),
-    col_types = cols_only(
-        `Sample Name` = col_character(),
-        `Session Name` = col_character(),
-        `EMG Peak-to-peak 1` = col_number()
-    )
-)
-
-View(t)
-View(get_subject_data(df, 9))
-df
+#df <- load_data() %>% clean_data()
+#write_csv(df, "df.csv")
+#View(df)
+#t <- read_delim("SWIP_018_S1.txt",
+#    delim = "\t", comment = "#", na = c("", "NA", "(null)"),
+#    col_types = cols_only(
+#        `Sample Name` = col_character(),
+#        `Session Name` = col_character(),
+#        `EMG Peak-to-peak 1` = col_number()
+#    )
+#)
+#
+#View(t)
+#View(get_subject_data(df, 9))
+#df

### 2023-06-11 jl/analysis.jl

@@ -35,23 +35,15 @@ function group_analysis(an, col=:SRA)
     rename!(an, 3 => :SRA)
 end
 
-df = CSV.read("LnDF.csv", DataFrame)
+df = CSV.read("df.csv", DataFrame)
 display(df)    
 
-s_df = subject_analysis(df)
-display(s_df)
-g_df = group_analysis(s_df)
-display(df)
+length(df.Sample)
+
+length(unique(df.Subject))
+print(unique(df.ISI))
+
+x = filter([:Subject, :ISI] => (x, y) -> x == 18 && y == 15, df)
+print(x)
 
 
-set_ρ(df)
-set_ρ_ln(df)
-set_ρ₂(df)
-set_ρ₂_ln(df)
-df = compute_inverse_variance_weights(df)
-set_weighted_ρ()
-set_weighted_ρ₂()
-
-display(df)
-
-CSV.write("full_df.csv", df)

### 2023-06-11 jl/main.jl

@@ -1,0 +1,47 @@
+using DataFrames
+using CSV
+using Statistics
+using Plots
+using StatsPlots
+using LinearAlgebra
+include("jl/math.jl")
+
+df = CSV.read("df.csv", DataFrame)
+
+function x_matrix(df::DataFrame, session_type::String)
+    """Computes matrix k x n matrix X as defined in the research notes. X is 
+    such that the row vectors r1, ..., rk are the paired responses of each of the 
+    k experimental subjects."""
+
+    filtered = filter([:ISI] => x -> x != -1, df)
+    k = length(unique(df.Subject))
+    X = Matrix{Float64}(undef, k, 192)
+
+    for i in unique(df.Subject)
+        subdf = filter([:Subject, :Session] => (x, y) -> x == i && y == session_type, filtered)
+        if length(subdf.EMGPeakToPeak) != 192
+            continue 
+        end
+        X = [X;transpose(subdf.EMGPeakToPeak)]
+    end
+    return X
+end
+
+function m_matrix(df::DataFrame, session_type::String, mprime::Bool=false)
+    filtered = filter([:ISI, :Type] => (x, y) -> x == -1 && y == session_type, df)
+    means::Vector{Float64} = []
+    for i in unique(df.Subject)
+        test_responses = filter([:Subject] => x -> x == i, filtered).EMGPeakToPeak
+        if mprime 
+            test_responses = test_responses.^(-1)
+        end
+        μ = mean(test_responses)
+        push!(means, μ)
+    end
+    Diagonal(means)
+end
+
+X = x_matrix(df, "SWD")
+M = m_matrix(df, "SWD")
+
+M * X

### 2023-06-11 jl/viewer.jl

@@ -1,0 +1,14 @@
+
+using DataFrames
+using CSV
+using Statistics
+using Plots
+using StatsPlots
+
+df2 = CSV.read("/data/SWIP_009_S1.txt", delim='\')
+
+test = filter(row -> row.Subject == 9 && row.Session == 1, df)
+display(test)
+
+plot(test.EMGPeakToPeak)
+

### 2023-06-12 paper/paper.tex
.

@@
 
-\begin{remark} 
+\begin{theorem} 
     $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
     property, consult the appendix.) 
-\end{remark}
+\end{theorem}
 
@@
 
-\begin{remark}
+\begin{theorem}
     Let $\pi$ denote the coefficient traditionally used for estimating the
@@
     (For a proof, consult the appendix.)
-\end{remark}
+\end{theorem}
 
@@
 
-\subsection{Generalizing to many subjects}
+\subsection{Generalizing to multiple subjects}
 
 \begin{definition}
-    Let $\textbf{x} \in \mathbb{R}^j$. Then we define $T(\textbf{x}) := \begin{bmatrix} 
+    Let $\textbf{x} \in \mathbb{R}^j$ with $j \in \mathbb{N}$. Then $T:
+    \mathbb{R}^j \to \mathbb{R}^j, R : \mathbb{R}^j \to \mathbb{R}^j$ are the
+    linear transformations defined as
+
+    $T(\textbf{x}) := \begin{bmatrix} 
         \rho(x_1) & \ldots & \rho(x_j)
@@
 
-\begin{theorem}
-    The mappings $T, R$ are linear transformations.
-\end{theorem}
-
-Let $k \in \mathbb{N}$ denote the number of experimental subjects in an
-experiment, to each of whom $n, m$ paired and test pulses were elicited,
+It is trivial to show that $T$ and $R$ are linear mappings (consult the
+Appendix). Now let $k \in \mathbb{N}$ denote the number of experimental subjects
+in an experiment, to each of whom $n, m$ paired and test pulses were elicited,
 respectively.
@@
 \begin{definition}
-    Let $\mu^{(t)}_i$ be average test-pulse response of the $i$th subject. Then
-    we define
-    $\textbf{M}^{k \times k}$ a diagonal matrix with $\textbf{M}_{(ii)} =
-    \frac{1}{\mu^{(t)}_{i}}$.
+    Let $\mu^{(t)}_i$ be average test-pulse response of the $i$th subject. Let
+    $\mu^{(t)^{-1}}_i$ be the average \textit{reciprocal} test-pulse response of
+    the $i$th subject. Then $\textbf{M}^{k \times k}$ is a diagonal matrix with
+    $\textbf{M}_{(ii)} := \frac{1}{\mu^{(t)}_{i}}$, and $\textbf{M'}^{~ k \times
+    k}$ is a diagonal matrix with $\textbf{M'}_{ii} := \mu^{(t^{-1})}_i$.
 \end{definition}
@@
 \begin{theorem}
-    The row-vectors of the matrix $\textbf{X}_\rho = \textbf{M}\textbf{P}$ are the
-    vectors $T(\textbf{x}^{(1)}), \ldots, T(\textbf{x}^{(k)})$.
-\end{theorem}
-
-Similarly, if we let $\textbf{M'}$ be a diagonal matrix such that
-$\textbf{M'}_{ii} = \mu^{(t^{-1})}_i$, where $\mu^{(t^{-1})}_i$ denotes the
-average \textit{inverse} test-pulse response of the $i$th subject, we come to
-the following result.
-
-\begin{theorem}
-    The row-vectors of the matrix $\textbf{X}_\delta = \textbf{M'}\textbf{P}$ 
+    The row-vectors of $\textbf{X}_\rho = \textbf{M}\textbf{P}$ are the
+    vectors $T(\textbf{x}^{(1)}), \ldots, T(\textbf{x}^{(k)})$, and
+    the row-vectors of $\textbf{X}_\delta = \textbf{M'}\textbf{P}$ 
     are the vectors $R(\textbf{x}^{(1)}), \ldots,
@@
 more than serving the algorithmic implementation of the model, which may
-prescind from them, these matrices provide insight into the nature of the
+prescind from them, these matrices provide insight into the exact nature of the
 transformations involved in the model.

### 2023-06-12 jl/loader.jl

@@ -1,0 +1,73 @@
+using CSV 
+using DataFrames 
+using DelimitedFiles 
+using "misc.jl"
+
+# SWIP_052_S1 has EMG Peak to peak 1 and 2. Ask Jen.
+
+const ISIORD120 = [
+    10, 0, 8, 0, -1, 10, 0, -1, 20, 0, -1, 15, 0, 10,
+    0, 20, 0, 8, 0, 10, 0, 15, 0, -1, 8, 0, -1, 10,
+    0, -1, 5, 0, -1, 5, 0, 10, 0, 8, 0, 4, 0, 8, 0,
+    -1, 4, 0, -1, 4, 0, 5, 0, -1, 20, 0, -1, -1, -1,
+    5, 0, 5, 0, 10, 0, -1, -1, -1, 4, 0, -1, 20, 0, 8,
+    0, 8, 0, 5, 0, -1, 15, 0, -1, 15, 0, -1, 5, 0, 15,
+    0, -1, 15, 0, 15, 0, 4, 0, 8, 0, 20, 0, 4, 0, 20,
+    0, 4, 0, -1, -1, 10, 0, 20, 0, -1, 15, 0, 5, 0, 20, 0, 4, 0
+]
+
+const ISIORDER72 = [
+    10, 8, -1, 10, -1, 20, -1, 15, 10, 20, 8, 10, 15, -1, 8, -1, 10, -1,
+    5, -1, 5, 10, 8, 4, 8, -1, 4, -1, 4, 5, -1, 20, -1, -1, -1, 5, 5,
+    10, -1, -1, -1, 4, -1, 20, 8, 8, 5, -1, 15, -1, 15, -1, 5, 15, -1,
+    15, 15, 4, 8, 20, 4, 20, 4, -1, -1, 10, 20, -1, 15, 5, 20, 4
+]
+
+
+function read_swipf(f::String)::Matrix
+    a = readdlm(f, '\t')
+    if size(a, 2) != 14
+        @warn "Inconsistent feature count on" f
+    end
+    a = a[:, setdiff(1:end, collect(2:size(a, 2) - 1))]
+    a[:, 1] = map(uppercase, a[:, 1])
+    from = findfirst(x -> occursin("MEP", x) || occursin("EVP", x), a[:, 1])
+    to = findlast(x -> occursin("END", x) || occursin("LAST", x), a[:, 1])
+    a = a[from:to, :]
+    a = a[:, setdiff(1:end, collect(2:size(a, 2) - 1))]
[…]

### 2023-06-12 misc.jl

@@ -1,0 +1,8 @@
+
+
+function str_to_num(s::String)
+    subject_pattern = r"[^0-9.-]"
+    eachmatch(subject_pattern, s)
+end
+
+"SWIP_009_S1.txt"[11]

### 2023-06-12 R/deprecated_loader.R
.

@@ -161,7 +161,7 @@ create_isi_col <- function(df) {
     if (len == 120) # If this ICF has 120 pulses.
         {
             vec <- ISI_ORDER_120
-        } else if (len == 72) # If this ICF has 120 pulses.
+        } else if (len == 72) # If this ICF has 72 pulses.
         {
             vec <- ISI_ORDER_72
         } else if (len == 118) {

### 2023-06-12 jl/loader.jl

@@ -1,7 +1,6 @@
 using CSV 
 using DataFrames 
 using DelimitedFiles 
-using "misc.jl"
 
 # SWIP_052_S1 has EMG Peak to peak 1 and 2. Ask Jen.
 
@@ -23,8 +22,45 @@ const ISIORDER72 = [
     15, 15, 4, 8, 20, 4, 20, 4, -1, -1, 10, 20, -1, 15, 5, 20, 4
 ]
 
+const PTINFO = CSV.read("TMS_ptinfo.csv", DataFrame)
+const COLNAMES = ["Sample", "Subject", "Group", "Session", "EMGPeakToPeak"]
 
+
+"""
+    read_swipf : String ↦ Matrix 
+
+    Read a .txt tab-delimited file with SWIP data as exported from BrainSight,
+    remove unwanted columns, and add categorical features.
+    
+    # Arguments
+    - `f::String`: The name of the file to read
+    
+    # Examples
+    ```julia-repl
+    julia> read_swipf("Data/SWIP_052_S1.txt")
+
+    93×7 Matrix{Any}:
+     "MEP_FIRST"   70  1  2  "BL"    nothing   232.593
+     "SAMPLE 15"   70  1  2  "BL"    nothing    95.7737
+     "SAMPLE 16"   70  1  2  "BL"    nothing    72.9705
+     "SAMPLE 17"   70  1  2  "BL"    nothing   201.809
+     "SAMPLE 18"   70  1  2  "BL"    nothing    77.5311
+     "SAMPLE 19"   70  1  2  "BL"    nothing   367.133
+     "SAMPLE 20"   70  1  2  "BL"    nothing   140.24
+     "SAMPLE 21"   70  1  2  "BL"    nothing   285.041
+     "SAMPLE 22"   70  1  2  "BL"    nothing   205.229
+     ⋮                                       ⋮         
+     "SAMPLE 99"   70  1  2  "BL"  -1          339.769
+     "SAMPLE 100"  70  1  2  "BL"  10         1014.75
+     "SAMPLE 101"  70  1  2  "BL"  20          281.62
+     "SAMPLE 102"  70  1  2  "BL"  -1          287.321
+     "SAMPLE 103"  70  1  2  "BL"  15          358.011
+     "SAMPLE 104"  70  1  2  "BL"   5           92.3532
+     "SAMPLE 105"  70  1  2  "BL"  20          107.175
+     "ICF_LAST"    70  1  2  "BL"   4           92.3532
[…]

### 2023-06-15 paper/paper.tex
.

@@
 
-    Let $x \in \textbf{x}$ be the evoked potential of a single paired
-    stimulation. Then we define two pulse-specific relative amplitude measures,
-    $\rho : \mathbb{R} \to \mathbb{R}, \delta : \mathbb{R} \to \mathbb{R}$,
-    where
+    Let $x \in \mathbb{R}^+$ be the evoked potential of a single paired
+    stimulation, and $\textbf{t} \in \mathbb{R}^m$. Then we define two
+    pulse-specific relative amplitude measures, $\rho : \mathbb{R} \times
+    \mathbb{R}^m \to
+    \mathbb{R}, \delta : \mathbb{R}\times \mathbb{R}^m \to \mathbb{R}$, where
 
     \begin{equation} 
-        \rho(x) := \frac{mx}{\sum_{i=1}^mt_j}
+        \rho(x, \textbf{t}) := \frac{mx}{\sum_{i=1}^mt_j}
     \end{equation}
@@
     \begin{equation} 
-        \delta(x) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} 
+        \delta(x, \textbf{t}) := \frac{x}{m}\sum_{j=1}^m\frac{1}{t_j} 
     \end{equation}
@@
 \begin{theorem} 
-    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
-    property, consult the appendix.) 
+    $\forall x: x \in \mathbb{R}^+:\delta(x, \textbf{t}) \geq \rho(x,
+    \textbf{t})$. (For a proof of this property, consult the appendix.) 
 \end{theorem}
 
-Notice that $\rho(x)$ is the proportion between the potential $x$, evoked by a
-paired stimulation, with respect to the average potential of single test
-stimulations. On the other hand, $\delta(x)$ is the average proportion of $x$
+The function $\rho$ represents the proportion between a potential $x$, evoked by
+a paired stimulation, with respect to the average potential of single test
+stimulations. On the other hand, $\delta$ is the average proportion of $x$
 with respect to each single test pulse.
@@
     \begin{align*}
-        \pi = \frac{\rho(x_1) + \ldots + \rho(x_n)}{n}
+        \pi = \frac{\rho(x_1, \textbf{t}) + \ldots + \rho(x_n, \textbf{t})}{n}
     .\end{align*}
@@
 neuroscientists have implicitly used it all along.
+
+\begin{remark}
+    Observe that 
+    \begin{align*}
+        \frac{\partial \rho}{\partial x } &= \frac{m}{\sum_{i = 1}^m
+        t_j} \\ \frac{\partial \rho}{\partial t_i} &= -\frac{xm}{\big( \sum_{j =
+        1}^m t_j \big)^2} \\ 
+                    \frac{\partial \delta}{\partial x} &= \frac{1}{m}\sum_{i =
+                    1}^m t_i \\ 
+                        \frac{\partial \delta}{\partial t_i} &= -\frac{x}{mt_i^2}
+    \end{align*}
+\end{remark}
+
+Since $\textbf{t}$ remains constant in the computation of the different relative
+amplitudes of $\textbf{x}_1, \ldots, \textbf{x}_n$, we shall use from now
+onwards the notation $\rho(x), \delta(x)$ to denote $\rho(x, \textbf{t})$ and
+$\delta(x, \textbf{t})$, respectively.
 

### 2023-06-15 jl/analysis.jl

@@ -36,6 +36,7 @@ function group_analysis(an, col=:SRA)
 end
 
 df = CSV.read("df.csv", DataFrame)
+df = X
 display(df)    
 
 length(df.Sample)

### 2023-06-15 jl/loader.jl

@@ -1,6 +1,6 @@
-using CSV 
-using DataFrames 
-using DelimitedFiles 
+using CSV
+using DataFrames
+using DelimitedFiles
 
 # SWIP_052_S1 has EMG Peak to peak 1 and 2. Ask Jen.
 
@@ -22,9 +22,55 @@ const ISIORDER72 = [
     15, 15, 4, 8, 20, 4, 20, 4, -1, -1, 10, 20, -1, 15, 5, 20, 4
 ]
 
-const PTINFO = CSV.read("TMS_ptinfo.csv", DataFrame)
-const COLNAMES = ["Sample", "Subject", "Group", "Session", "EMGPeakToPeak"]
+const S09_S1_BAD = [35, 37, 38, 40, 41, 43, 44, 46, 48, 50, 52, 54, 56, 57, 58, 59, 60,
+    62:68, 70:85, 87:94, 96:192, 104:111, 113, 114,
+    116:119, 121, 122, 124, 126:130, 132:134, 136:140,
+    142, 144, 145, 147, 148, 149, 151, 152]
 
+const S18_S1_GOOD = [81, 82, 84, 87, 89, 90, 93, 112,
+    124, 129, 131, 134, 136, 137, 140, 142, 147, 149, 151]
+const S18_S2_BAD = [38, 40, 42, 49, 51, 56, 59,
+    61:66, 68:72, 74:78, 82, 83,
+    85, 87, 88, 90, 93, 96, 98, 99,
+    100, 103, 105, 107] #107 is last icf
+
+const S49_S2_BAD = [55, 57, 116, 117, 120,
+    121, 124, 125, 134, 135,
+    136, 137, 152, 153, 154,
+    161, 162, 170, 171]
+
+const S53_S2_BADS = [50, 53, 56, 58, 61, #MEPS thus far
+    129, 155, 156] # ICFs
+
+const S76_BADS = [22, 23, 24, 25, 26, 27, 28, 31, 34, 38, 39, 40, 41]
+
+const S80_S1_BADS = [103, 104, 118, 119, 120, 134, 137, 139]
+const S80_S2_BADS = [36, 37, 103, 104, 38, 112, 113, 116, 118, 119, 120, 128, 136, 137, 152, 153]
+
+const S87_S2_BADS = [44, 45, 46, 47, 95, 97, 98, 102, 103,
+    106, 107, 126:129, 143, 168, 169, 172, 173, 181:184]
+
[…]

### 2023-07-05 paper/paper.tex
.

@@
 
-The function $\rho$ represents the proportion between a potential $x$, evoked by
-a paired stimulation, with respect to the average potential of single test
-stimulations. On the other hand, $\delta$ is the average proportion of $x$
-with respect to each single test pulse.
+The function $\rho$ represents the proportion between a potential $x \in
+\textbf{x}$ with respect to the average
+potential of single test stimulations. On the other hand, $\delta$ is the
+average proportion of $x$ with respect to each single test pulse.
 
@@
 \begin{theorem}
-    Let $\pi$ denote the coefficient traditionally used for estimating the
-    relative amplitude between test and paired pulses. In accordance to what was
-    described in the second paragraph of Section $4$,
+
+    Let
 
     \begin{align*}
-        \pi = \frac{\mu_x}{\mu_t}
-    .\end{align*}
+        \tau = \frac{\bar{x}}{\bar{t}}
+    \end{align*}
 
-    where $\mu_x, \mu_t$ are the mean paired and test pulse responses,
-    respectively. It follows from the definition of $\pi$ and $\rho$ that
+    where $\bar{x}, \bar{t}$ are the average paired and test pulse responses,
+    respectively. Then
 
     \begin{align*}
-        \pi = \frac{\rho(x_1, \textbf{t}) + \ldots + \rho(x_n, \textbf{t})}{n}
-    .\end{align*}
+        \tau = \frac{1}{n} \sum_{i = 1}^n \rho(x_i, \textbf{t})
+    \end{align*}
 
@@
 
-The second remark is of great importance. It states that the way in which
+Here, $\tau$ is the coefficient traditionally used for estimating the relative
+amplitude between test and paired pulses (Lefaucher, 2019), in accordance to
+what was described in the second paragraph of \textbf{Section 4}.
+
+The previous theorem is of great importance. It states that the way in which
 relative amplitude was traditionally computed---the way in which neuroplasticity
@@
 in terms of scale. More specifically, the traditional coefficient used to proxy
-neuroplasticity at the subject level is nothing but the mean $\rho$
-measure in an experimental session. We believe this bestows $\rho$ with a
-certain degree of confidence, insofar as not only it is not entirely new, but
-neuroscientists have implicitly used it all along.
-
-\begin{remark}
-    Observe that 
-    \begin{align*}
-        \frac{\partial \rho}{\partial x } &= \frac{m}{\sum_{i = 1}^m
-        t_j} \\ \frac{\partial \rho}{\partial t_i} &= -\frac{xm}{\big( \sum_{j =
-        1}^m t_j \big)^2} \\ 
-                    \frac{\partial \delta}{\partial x} &= \frac{1}{m}\sum_{i =
-                    1}^m t_i \\ 
-                        \frac{\partial \delta}{\partial t_i} &= -\frac{x}{mt_i^2}
-    \end{align*}
-\end{remark}
+neuroplasticity at the subject level is nothing but the mean $\rho$ measure in
+an experimental session. We believe this bestows $\rho$ with a certain degree of
+confidence, insofar as not only it is not entirely new, but neuroscientists have
+implicitly used it all along.
 
@@
 \begin{definition}
-    Let $\textbf{x} \in \mathbb{R}^j$ with $j \in \mathbb{N}$. Then $T:
-    \mathbb{R}^j \to \mathbb{R}^j, R : \mathbb{R}^j \to \mathbb{R}^j$ are the
+    Let $\textbf{x} \in \mathbb{R}^n$ with $n \in \mathbb{N}$. Then $T:
+    \mathbb{R}^n \to \mathbb{R}^n, R : \mathbb{R}^n \to \mathbb{R}^n$ are the
     linear transformations defined as
 
-    $T(\textbf{x}) := \begin{bmatrix} 
-        \rho(x_1) & \ldots & \rho(x_j)
-    \end{bmatrix}^\intercal $ and $R(\textbf{x}) := \begin{bmatrix} 
-        \delta(x_1) & \ldots & \delta(x_j) 
-    \end{bmatrix}^\intercal $.
+    \begin{align*}
+        T(\textbf{x}) &:= \begin{bmatrix} 
+        \rho(x_1) & \ldots & \rho(x_n)
+        \end{bmatrix}^\intercal \\ R(\textbf{x}) &:= \begin{bmatrix} 
+        \delta(x_1) & \ldots & \delta(x_n) 
+    \end{bmatrix}^\intercal
+    \end{align*}
+
 \end{definition}
@@
 \begin{definition}
-    Let $\mu^{(t)}_i$ be average test-pulse response of the $i$th subject. Let
-    $\mu^{(t)^{-1}}_i$ be the average \textit{reciprocal} test-pulse response of
-    the $i$th subject. Then $\textbf{M}^{k \times k}$ is a diagonal matrix with
-    $\textbf{M}_{(ii)} := \frac{1}{\mu^{(t)}_{i}}$, and $\textbf{M'}^{~ k \times
-    k}$ is a diagonal matrix with $\textbf{M'}_{ii} := \mu^{(t^{-1})}_i$.
+    Let $\overline{t}_i$ be average test-pulse response of the $i$th subject.
+    Let $\hat{t}_i$ be the average of the \textit{reciprocal}
+    responses on the $i$th subject. Then $\textbf{M}^{k \times k}$ is a diagonal
+    matrix with $\textbf{M}_{(ii)} := \frac{1}{\overline{t}_i}$, and
+    $\textbf{M'}^{~ k \times k}$ is a diagonal matrix with $\textbf{M'}_{ii} :=
+    \hat{t}_i$.
 \end{definition}
@@
 many ways to deal with this situation in the implementation of our model---the
-safest being the direct application of $T$ and $R$ to the data vectors---. Thus,
-more than serving the algorithmic implementation of the model, which may
-prescind from them, these matrices provide insight into the exact nature of the
-transformations involved in the model.
+safest being the direct application of $T$ and $R$ to the artifact-excluded data
+vectors---. Thus, more than serving the algorithmic implementation of the model,
+which may prescind from them, these matrices provide insight into the exact
+nature of the transformations involved in the model.
 
@@
 
+\subsection{Further theory}
+
+Observe that
+
+\begin{align*}
+    \frac{\partial \rho}{\partial x } &= \frac{1}{\overline{t}}  \\ 
+     \frac{\partial \rho}{\partial t_i} &= -\frac{x}{m\overline{t}^{~ 2}}
+\end{align*}
+
+We recall that $t_i, x_i \in \mathbb{R}^{+}$ by assumption, from which follows
+that $\rho$ is constantly increasing with respect to $x$ and
+decreasing with respect to $\textbf{t}$. Similarly,
+
+\begin{align*} 
+    \frac{\partial \delta}{\partial x} &= \overline{t}\\ 
+    \frac{\partial \delta}{\partial t_i} &= -\frac{x}{mt_i^2}
+\end{align*}
+
+This result extends the property noted in \textbf{Theorem 1}. Namely, not only
+is $\delta$ greater than $\rho$ for any $x$, but it also has a greater constant
+rate of change when $t > 1$---which is always the case with non-anomalous EMG
+peak to peak values.
+
+Furthermore, we found that $\rho$ can be alternatively defined as
+
+\begin{align*}
+    \delta(x, \textbf{t}) = -\frac{x}{m}\int \frac{1}{t_i^2} dt_i \\ 
+    \rho(x, \textbf{t}) &= -\frac{x}{m}\int \frac{1}{\overline{t}^2}
+\end{align*}
+
+\subsection{Proofs}
+
 \textbf{Proof 1}. In \textbf{Remark 1}, we observed the following property:
@@
 \begin{align*} 
-    \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\ 
-    S_2^mS_1^m &\geq m^2
+    &\frac{x}{m}\sum^m\frac{1}{t_j} \geq \frac{xm}{\sum^m t_j} \iff 
+    S_2^mS_1^m \geq m^2
 \end{align*}
@@
 
-We know $S^k_1S^k_2 \geq k^2$ and then it suffices to show $t_{k+1}S_1^k +
+By hypothesis, $S^k_1S^k_2 \geq k^2$, and then it suffices to show $t_{k+1}S_1^k +
 \frac{S^k_2}{t_{k+1}}\geq 2k$. To prove this, simply observe that
@@
 
-\textbf{Proof 2.} In \textbf{Remark 2}, we observe that $\rho$ is implicitly
+~
+
+\textbf{Proof 2.} In \textbf{Theorem 2}, we observed that $\rho$ is implicitly
 used in the traditional measures of relative amplitude. Recall that such methods
@@
 
-\begin{align*} \frac{\mu_x}{\mu_d} = \frac{\frac{1}{n}\sum_{j=1}^n
-        x_{j}}{\frac{1}{m}\sum_{j = 1}^m t_j} 
-        = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
+\begin{align*} 
+    \frac{\bar{x}}{\bar{t}}
+         &= \frac{1}{n} \Big(\frac{ \sum_{i = 1}^n mx_i}{\sum_{i = 1}^m
+             t_i}\Big) \\ 
+         &= \frac{1}{n} \Big(\frac{ mx_1}{\sum_{i = 1}^m t_i} + \ldots + \frac{
+             mx_n}{\sum_{i = 1}^m t_i}\Big) \\ 
+         &= \frac{1}{n} \sum_{i = 1}^n \rho(x_i, \textbf{t})
 \end{align*}
 
-\begin{proof}[Theorem 1]
+\begin{proof}[Theorem 3]
 
@@
 
-    The proof the $R$ is linear is practically identical.
+    The proof that $R$ is linear is practically identical.
 \end{proof}

### 2023-07-05 R/plots.R

@@ -1,10 +1,113 @@
-source("R/loader.R")
+source("r/loader.r")
 library(ggplot2)
+library(tidyr)
+library(readr)
+library(dplyr)
+library(tibble)
+library(robustbase)
+library(lmPerm)
+library(MASS)
+
+df <- read_csv("JLRA_DF.csv")
+df <- filter(df, ISI != 0)
+df$ISI <- as.factor(df$ISI)
+df$Group  <- as.factor(df$Group)
 
 plot_analysis <- function(df) {
-    p <- ggplot(df, aes(x = ISI, y = SRA, group = Label, color = Label)) +
+    p <- ggplot(df, aes(x = isi, y = sra, group = label, color = label)) +
         geom_line() +
-        geom_point(aes(x = ISI, y = SRA, group = Label)) +
+        geom_point(aes(x = isi, y = sra, group = label)) +
         geom_hline(yintercept = 1, linetype = "dashed")
     p
 }
+
+
+
+# ------------ 
+
+HCSWD <- df %>% filter(Label == "HC SWD") %>% filter(Subject != 8)
+HCBL <- df %>% filter(Label == "HC BL")
+MDDSWD <- df %>% filter(Label == "MDD SWD") %>% filter(Subject != 8)
+MDDBL <- df %>% filter(Label == "MDD BL")
+
+# Is median before SWD greater? 
+wilcox.test(HCBL$EMGPeakToPeak, HCSWD$EMGPeakToPeak, paired=TRUE, 
+            alternative="greater")
+# Is median after SWD lower?
+wilcox.test(MDDBL$EMGPeakToPeak, MDDSWD$EMGPeakToPeak, paired=TRUE, 
+            alternative="less")
+
+wilcox.test(HCBL$EMGPeakToPeak, MDDBL$EMGPeakToPeak,
+            alternative="greater")
+
+wilcox.test(HCBL$RRA, HCSWD$RRA, paired=TRUE)
+# Is median after SWD lower?
+wilcox.test(MDDBL$RRA, MDDSWD$RRA, paired=TRUE, alternative="less")
+
+wilcox.test(HCSWD$RRA, MDDSWD$RRA)
+
+library(coin)
+
+df
+X <- filter(df, Group == 1) %>% filter(Subject != 8)
+
+median_test(RRA~as.factor(Label), data=X)
+oneway_test(RRA~as.factor(Label),data=X)
+
+# Calculate mean and standard deviation by group using dplyr
+summary_stats <- df %>%
[…]

### 2023-07-05 jl/analysis.jl

@@ -3,6 +3,7 @@ using CSV
 using Statistics
 using Plots
 using StatsPlots
+using Clustering
 include("jl/math.jl")
 
 function sra(pulses, subject, session)
@@ -35,9 +36,25 @@ function group_analysis(an, col=:SRA)
     rename!(an, 3 => :SRA)
 end
 
-df = CSV.read("df.csv", DataFrame)
-df = X
+sdf = subject_analysis(df)
+gdf = group_analysis(sdf)
+
+grouped = groupby(gdf, [:Label, ISI])
+plot()
+x = gdf.ISI
+for group in grouped 
+    print(group)
+    cat = group[1, :Label][1]
+    y = group[:, :SRA]
+    plot!(x, y, label=cat, marker = :circle)
+end
+display(plot)
+
+df = CSV.read("full_df.csv", DataFrame)
+df = dropmissing(df)
+sdf = subject_analysis(df)
 display(df)    
+CSV.write("GROUPAN_DF.csv", gdf)
 
 length(df.Sample)
 
@@ -47,4 +64,8 @@ print(unique(df.ISI))
 x = filter([:Subject, :ISI] => (x, y) -> x == 18 && y == 15, df)
 print(x)
 
+features = hcat(df.ISI, df.EMGPeakToPeak, df.RRA, df.RRA2)
+result = kmeans(features, 4)
+df
+scatter(df.RRA, df.RRA2, marker_z = result.assignments)
 

### 2023-07-05 jl/loader.jl

@@ -22,18 +22,6 @@ const ISIORDER72 = [
     15, 15, 4, 8, 20, 4, 20, 4, -1, -1, 10, 20, -1, 15, 5, 20, 4
 ]
 
-const S09_S1_BAD = [35, 37, 38, 40, 41, 43, 44, 46, 48, 50, 52, 54, 56, 57, 58, 59, 60,
-    62:68, 70:85, 87:94, 96:192, 104:111, 113, 114,
-    116:119, 121, 122, 124, 126:130, 132:134, 136:140,
-    142, 144, 145, 147, 148, 149, 151, 152]
-
-const S18_S1_GOOD = [81, 82, 84, 87, 89, 90, 93, 112,
-    124, 129, 131, 134, 136, 137, 140, 142, 147, 149, 151]
-const S18_S2_BAD = [38, 40, 42, 49, 51, 56, 59,
-    61:66, 68:72, 74:78, 82, 83,
-    85, 87, 88, 90, 93, 96, 98, 99,
-    100, 103, 105, 107] #107 is last icf
-
 const S49_S2_BAD = [55, 57, 116, 117, 120,
     121, 124, 125, 134, 135,
     136, 137, 152, 153, 154,
@@ -106,7 +94,6 @@ const TOTAL_BADS = length(BADS)
      "ICF_LAST"    70  1  2  "BL"   4           92.3532
 """
 function read_swipf(f::String)::Matrix
-
     a = readdlm(f, '\t')
     if size(a, 2) != 14
         @warn "Inconsistent feature count on" f
@@ -164,7 +151,7 @@ function add_isi_feature(X::Matrix, f::String)
     icf_count = to - from + 1
     if icf_count != 120 && icf_count != 72
         print(f, "\n")
-        @warn "WARNING: Erroneous ICF count on " f
+        @warn "WARNING: Erroneous ICF count on " f " with count " icf_count
     end
     if any(x -> !isa(x, Real), X[:, 2])
         error("Non-numeric EMG ", f)
@@ -238,11 +225,3 @@ function get_label(group, type)::String
         return type == "BL" ? "MDD BL" : "MDD SWD"
     end
 end
-
-get_label.(X.Group, X.Type)
-
-read_swipf("Data/SWIP_070_S2.txt")
-X = read_swipf()
-
-rename!(X, COLNAMES)
-CSV.write("JL_DF.csv", X)

### 2023-07-05 jl/tests.jl

@@ -1,0 +1,158 @@
+using HypothesisTests
+using CSV 
+using DataFrames
+using EffectSizes
+using Statistics
+using Plots
+using StatsPlots
+
+df = CSV.read("JLRA_DF.csv", DataFrame)
+df = filter(:ISI => x -> x != 0, df)
+
+function label_filter(label::String, df)
+    filter([:Label, :Subject] => (x, y) -> x == label && 
+                                                y != 8, 
+                                                df)
+end
+
+function run_dependent_sample_tests(df::DataFrame)
+    HC = [label_filter("HC BL", df), label_filter("HC SWD", df), "HC"]
+    MDD = [label_filter("MDD BL", df), label_filter("MDD SWD", df), "MDD"]
+
+    for group in [HC, MDD]
+        println("~~~~~~~~~~~~~~~~~~~~~~  ", group[3], "  ~~~~~~~~~~~~~~~~~~~~~~~")
+        println(SignTest(group[1].EMGPeakToPeak, group[2].EMGPeakToPeak))
+        cohen_display(CohenD(group[1].EMGPeakToPeak, group[2].EMGPeakToPeak))
+    end
+end
+
+function run_independent_sample_tests(df::DataFrame)
+    A = [label_filter("HC BL", df), label_filter("MDD BL", df), "HC BL vs MDD BL"]
+    B = [label_filter("HC SWD", df), label_filter("MDD SWD", df), "HC SWD vs MDD SWD"]
+    display(A[1])
+    display(B[1])
+
+    for group in [A, B]
+        println("~~~~~~~~~~~~~~~~~~~~~~  ", group[3], "  ~~~~~~~~~~~~~~~~~~~~~~~")
+        println(MannWhitneyUTest(group[1].EMGPeakToPeak, group[2].EMGPeakToPeak))
+    end
+end
+
+run_independent_sample_tests(mepdf)
+
+function run_tests(df::DataFrame)
+    HC_BL = label_filter("HC BL", df); HC_SWD = label_filter("HC SWD", df)
+    MDD_BL = label_filter("MDD BL", df); MDD_SWD = label_filter("MDD SWD", df)
+   
+    println("----------------------------------------------------")
+    println("Running tests on independent samples ~ HCs vs. MDDs Baseline")
+    println("----------------------------------------------------\n\n")
[…]
